The structure-interaction model of polymyxin lipopeptides with human oligopeptide transporter 2

Version 1

Abstract

Background

Multidrug-resistant (MDR) Gram-negative bacteria have triggered a critical global health crisis. Polymyxin lipopeptide antibiotics are used as a last-line therapy against these problematic pathogens, but their clinical use is largely limited by severe nephrotoxicity. Human oligopeptide transporter 2 (hPepT2) is a membrane transporter mediating the reabsorption of polymyxins in renal proximal tubular cells, substantially contributing to their nephrotoxicity. However, it remains unclear how polymyxins interact with hPepT2.

Methods

In this study, we investigated the structure-interaction relationship (SIR) of polymyxins with hPepT2 by integrating computational, chemical and cell biology approaches. Bioinformatic modelling predicted the residues essential for the binding of polymyxins with hPepT2. Transporter mutagenesis and molecular analysis were employed to explore the role of each residue in the interaction of hPepT2 and polymyxins. Moreover, we synthesised a series of polymyxin-like analogues with altering the moieties that are critical for binding with hPepT2. The antibacterial activity and nephrotoxicity of these analogues were subsequently assessed.

Results

Our bioinformatic modelling proposed an outward-facing structure of hPepT2 with a possible transport pathway that polymyxins bind to the lateral opening site of hPepT2 (e.g. E214, D215, D317, D342, E622). Molecular assays for transporter function and expression confirmed that D215 residue of hPepT2 is critical for polymyxin binding, while several other residues significantly impact on transporter turnover rate and/or protein expression. Our experimental validations showed that the lipopeptide analogues with altering the Dab1, Dab3, Dab5 and Dab9 moieties of polymyxins demonstrated decreased interactions with hPepT2. Among these synthetic analogues, alanine substitution at Dab3 showed reduced nephrotoxicity in mice while reserved antibacterial activity against a range of bacterial strains.

Conclusions

Overall, this proof-of-concept study demonstrated that the computationally predicted and experimentally validated polymyxin-hPepT2 SIR model provides a viable approach for the discovery of novel, safer lipopeptide antibiotics.

Introduction

The World Health Organization has highlighted antimicrobial resistance as an urgent threat to global public health (1). In particular, the Gram-negative ‘superbugs’ Acinetobacter baumannii, Pseudomonas aeruginosa, and Klebsiella pneumoniae can develop resistance to nearly all available antibiotics, posing some of the most serious challenges in modern medicine (2). Due to the lack of effective alternative antibiotics, polymyxins are often used as a last-line therapy against these ‘superbugs’ (3). However, polymyxins can cause nephrotoxicity in up to 60% of patients after intravenous administration, which is the major dose-limiting factor for their clinical use (4, 5). Thus, there is an urgent need for safer new-generation polymyxin lipopeptide antibiotics (6).

It has been demonstrated that polymyxin-induced nephrotoxicity is associated with extensive reabsorption into renal proximal tubular cells (4). The intracellular concentration of polymyxins in renal proximal tubular cells can be 5,000-fold higher than their extracellular concentration, which consequently leads to oxidative stress, autophagy, cell cycle arrest, and apoptosis (4). Therefore, the extensive reabsorption of polymyxins in renal proximal tubular cells is the very first step of nephrotoxicity.

Human Solute Carrier transporters (SLCs) are the primary influx transporters responsible for the cellular uptake of many drugs, including antibiotics (7, 8). There are over 300 SLC isomembers, among which SLCO, SLC22A and SLC15A are the key subfamilies mediating drugs moving across cell membrane (9, 10). Our previous study revealed that the cellular uptake of polymyxins is mediated through hPepT2 in the kidney (11), which is encoded by a key isoform of SLC15A subfamily, human oligopeptide transporter 2 (hPepT2) (10). This transporter protein is highly expressed in renal tubular cells mediating the uptake of polymyxins, with a comparable kinetic constant (Km, reflecting transporter-substrate binding affinity) to its prototype substrates (10–12). It is one of the proton-driven symporters using the inwardly direct proton electrochemical gradient to drive the concentrative uptake of substrates across cell membrane (13). Substrates of hPepT2 include a range of di- or tri-peptides and peptide-like drugs, such as angiotensin-converting enzyme inhibitors (14). However, it remains unclear how hPepT2 recognise lipopeptide antibiotics, particularly polymyxins, which is critical for understanding its interaction with polymyxins contributing to their nephrotoxicity.

Here, we integrated computational and experimental approaches to elucidate the interaction relationship between hPepT2 and polymyxins. Importantly, we conducted a proof-of-concept chemical biological study by designing novel lipopeptides and evaluated their cellular uptake, nephrotoxicity, and antibacterial activities. This study establishes the first structure-interaction relationship (SIR) model for polymyxins and hPepT2, laying the foundation for rational drug design and discovery of novel lipopeptide antibiotics.

Results

MD prediction of the structural model of outward-facing hPepT2

We developed the three-dimensional structure of hPepT2 using AlphaFold2, as the outward structural model of hPepT2 (Fig. 1a) has not been experimentally solved. The putative overall structure adopted an inward open conformation (Fig. 1b), which disabled the exploration of the initial capture of polymyxins from the extracellular side of the membrane. Parker et al. demonstrated the Cryo-electron microscopy (Cryo-EM) structure of rat PepT2 with an outward open conformation (15). As rat PepT2 shares 83% sequence similarity with hPepT2, it is desired to be the prototype in the exploration of hPepT2 structure through the homologous modelling algorithm. Upon remodelling, the resultant structure of hPepT2 adopted an outward open conformation (Fig. 1c). Several transmembrane domains (TMs) are involved in the formation of the solvent-accessible cavity for polymyxin translocation, including TM1, TM2, TM5, TM7, TM8 and TM10. The gate residue pairs (L75-R329, P341-E79, D170-S635 and K642-A177) marking the conformations were clearly indicated in the solvent-accessible regions on both the extracellular and intracellular surfaces (Fig. 1d). During 500-ns all-atom MD simulations with the inward-facing conformation as an initial configuration, the distances of the extracellular gate residue pairs L75-R329 and P341-E79 increased from 20 Å to 31 Å, and from 15 Å to 21 Å, respectively. Meanwhile, the distances of the intracellular gate residue pairs D170-S635 and K642-A177 decreased from 15 Å to 13 Å, and from 13 Å to 10 Å, respectively (Fig. 1e). Noteworthy, the conformational transition from the inward pose to the outward pose uncovered by MD simulations is consistent with the homologous model of rat PepT2 regarding the distance of extracellular and intracellular gate (15). This observation indicates the accuracy of the proposed model of hPepT2.

Structural conformations of hPepT2.

(a) The structural toplogy of hPepT2. (b) The structural model of hPepT2 built with the AlphaFold approach. (c) The structural model of hPepT2 built with the homologous modeling method. The three dimensional structure of rat PepT2 (PDB code: 7NQK) is used as the structural template. (d) The residues that are used to define the geometry of the channel of hPepT2. (e) The distances between the selected residues and representative structural snapshots are analyzed based on MD simulations.

Polymyxin B binds to the lateral opening site of hPepT2 to enter the transport pathway

To further explore how hPepT2 recognises polymyxin molecules, we examined their interaction through long-time scale coarse-grained MD simulations. Although ten polymyxin B molecules were introduced to the simulation system, only one polymyxin molecule was observed to bind to the entrance of hPepT2 binding pocket, while the other polymyxin molecules attached to the membrane lipids (Fig. S3). Importantly, we discovered that there is a distinct pathway that polymyxins reached the extracellular lateral opening gate of hPepT2, as revealed by the simulation replicates (Fig. 2a and S4) with distinct equilibrated binding conformations (Fig. 2b). Our MD simulations predicted a possible transport pathway for polymyxins via hPepT2. Initially, polymyxin B attached to the plateau of TM1-TM6 helix bundle, mainly driven by the electrostatic interactions between the Dab5, Dab8 and Dab9 residues of polymyxin B and the E79, D208, E214 and D215 residues of hPepT2 (Fig. 3). Subsequently, polymyxin B translocated to the junction between the TM7-TM12 helix bundles and extracellular domains (ECDs) of hPepT2, where the Dab1, Dab3 and Dab5 residues of polymyxin B electrostatically interacted with the D342 and E555 residues of hPepT2. Polymyxin B then stabilised at the lateral opening gate of hPepT2 (i.e. lateral entry site), where the Dab8 and Dab9 moieties of polymyxin B electrostatically interacted with the D342 and E555 residues of hPepT2. Moreover, the hydrophobic atoms in Dab8 and Dab9 interacted with the F200, I201 and M204 residues of hPepT2, while the fatty acyl group of polymyxin B bound with the hydrocarbon tails of the phospholipid bilayer of the membrane.

Spontaneous interactions of polymyxin B with hPepT2.

(a) The interaction pathway of polymyxin B with hPepT2. The spatial locations of polymyxin B molecules are represented with colored spheres. The color spectrum from cyan to red indicates the simulation continuity from 0 to 3 μs. (b) The equilibrated conformation of hPepT2-polymyxin complex according to the proposed pathway. The polymyxin B molecule is shown in brown and blue spheres with its D-Phe6 and N-terminal fatty acyl group labeled.

Key structural moieties and residues involved in the interaction of polymyxin B with hPepT2.

The surface potential of hPepT2 is shown with red color indicating negative potential and blue color indicating positive potential.

To examine the detailed interactions between hPepT2 and polymyxin B, we performed all-atom MD simulations of the hPepT2-polymyxin B complex. During system preparation, the polymyxin B molecule was placed in four different poses within the complex to enhance sampling of binding dynamics. The results revealed that Dab1, Dab8, and Dab9 form strong electrostatic interactions with hPepT2, with interaction energies ranging from – 180 to –260 kJ/mol (Fig. 4), while Dab3 and Dab5 may also contribute significantly. Together with observations from the coarse-grained simulations, these results indicate that the positively charged Dab residues are key determinants of binding affinity to hPepT2.

Interaction energy between hPepT2 and each residue of polymyxin B.

The interaction energy was calculated using four MD simulation replicates. Note: Coul: electrostatic interaction; VDW: hydrophobic interaction.

Functional mutagenesis elaborates on the predicted transport pathway

At physiological pH, there are five positively charged residues (i.e. Dab residues) on the polymyxin molecule and we assumed that the key negatively charged residues of hPepT2 identified by the MD simulations (i.e. E79, D208, E214, D215, D317, D342, E555 and E622) may play crucial roles in the recognition and interactions with polymyxins. We employed functional mutagenesis to confirm their roles in the uptake of polymyxins, as predicted by our MD studies. Through the commonly used alanine-scanning mutagenesis approach, we constructed the correspondent mutants of hPepT2. Noteworthy, alanine substitution not only impacted on the side chain length of the proposed residues but also altered their charges from negative to neutral. It is postulated that such a drastic change to the residues of in hPepT2 may largely affect their interactions with polymyxins.

The transport function of these hPepT2 mutants were evaluated as to the uptake of H3-Gly-Sar (a typical substrate of hPepT2) (11) as well as that of the fluorescent polymyxin probe MIPS-9541 (Fig. 5). The fluorescence images were also captured as the qualitative measurement of the overexpressing cells exposed to our dansyl-fluorescent polymyxin probe MIPS-9541, which indicated the consistent results (Fig. 6). Our data showed that E79A, D80A and D208A mutants maintained the transport activity in the uptake of both substrates. In contrast, E214A, D215A, D342A and E622A mutants had impaired function in the transport of either substrate. Interestingly, D317A and E555A showed differential activity in the uptake of these substrates, suggesting the distinct roles of both residues in the interaction with different substrates.

Uptake of 3H-Glycosarcosine and polymyxin B fluorescence probe MIPS-9541 by hPepT2 and its mutants.

HEK293 cells were transfected with hPepT2 and its mutant constructs. (a) Uptake of 3H-Gly-Sar (5 µM) by hPepT2 or its mutants. (b) Uptake of MIPS-9541 (10 µM) by hPepT2 or its mutants. ***p < 0.001 vs. control by Welch’s t-test.

Uptake of polymyxin B fluorescence probe MIPS-9541 in HEK293 cells overexpressing hPepT2 or its mutants by fluorescence imaging.

Note: WT: Wild type; NC: Negative control.

Subsequently, kinetic analysis was conducted to characterise those hPepT2 mutants with impaired activity in the uptake of MIPS-9541 (i.e. E214A, D215A, D342A and E555A). Noteworthy, the E622A mutant was not included in this analysis due to its critically low transporter function. As shown in Table 1, the D215A mutant demonstrated a doubled Km (96.6 ± 20.9 µM of D215A vs. 47.1 ± 11.9 µM of the wild type) and Vmax (8610 ± 1363 pmol/(µg/4 min) of D215A vs. 4252 ± 652.9 pmol/(µg/4 min) of the wild type). It was also evident that the D342A mutant showed a mildly increased Km (66.9 ± 13.2 µM of D342A vs. 47.1 ± 11.9 µM of the wild type) without altering its Vmax. The Vmax values of E214A and E555A were significantly reduced (2332 ± 24.2 pmol/(µg/4 min) of E214A and 2376 ± 320.8 pmol/(µg/4 min) of E555A vs. 4252 ± 652.9 pmol/(µg/4 min) of the wild type). However, the Km values of E214A and E555A mutants were moderately decreased compared to that of the wild type (24.4 ± 4.48 µM of E214A and 26.7 ± 6.99 µM of E555A vs. 47.1 ± 11.9 µM of wild type).

Noteworthy, a change of Vmax value may be due to altered transporter protein expression and/or transporter turnover rate. Thus, the total cell and plasma membrane expression of the hPepT2 and its mutants were investigated to further examine the molecular mechanism underpinning the functional alternations of these mutants. As shown in Fig. 7, E214A, D215A, E555A and E622A significantly reduced total cell expression and cell surface expression. In the case of D342A, it retained the transporter protein expression in the whole cell but had a markedly impaired cell surface expression.

Total cell and cell surface expression of hPepT2-c-Flag and its mutants.

HEK293 cells were transfected with hPepT2-c-Flag and its mutant constructs. (a) Cells were lysed and subject to SDS-electrophoresis. Immunoblots were probed with anti-Flag antibody and then followed by anti-actin antibody. Representative images of each hPepT2 mutant are shown. (b) Cell surface proteins were labelled with NHS-ss-Biotin and then pulled down with Streptavidin-agarose beads. Biotin-labelled surface protein samples were subject to SDS-electrophoresis and the immunoblots were probed with anti-Flag antibody. Equal portion of the supernatant samples were separated by SDS-electrophoresis and the immunoblots were incubated with anti-actin antibody. Representative images of each hPeT2 mutant are shown. (c) Densitometry analysis of the relative total cell expression of hPepT2 mutants (ratios of Flag/actin). (d) Densitometry analysis of the relative cell surface expression of hPepT2 mutants (ratios of Flag/actin). Data are presented as % of the hPepT2 wildtype control (mean ± SD). Experiments were repeated on three occasions. *p < 0.05; **p < 0.01 vs. wildtype control by unpaired t-test. Note: WT: wild type; NC: vector-transfected negative control.

Chemical biology of polymyxin interaction with hPepT2

As described above, the Dab1, Dab3, Dab5, Dab8 and Dab9 residues of polymyxin B were predicted to be involved with its interaction with hPepT2. To test this hypothesis, we designed and synthesised five polymyxin B analogues with substitutions of Alanine on these Dab positions. We performed transport uptake assay to evaluate their interactions with hPepT2. As shown in Fig. 8, the lipopeptide analogues of FADDI-170, FADDI-175, FADDI-793 and FADDI-795 but not FADDI-167 showed reduced uptake in hPepT2 expressing cells, indicating that Dab1, Dab3, Dab5 and Dab9 but not Dab8 are likely involved in the interaction with hPepT2. Therefore, these four lipopeptide analogues may have less intracellular accumulation and reduced nephrotoxicity.

Uptake of polymyxin B fluorescence probe MIPS-9541 and polymyxin analogues by hPepT2.

The data are expressed as the fold of the uptake via hPepT2 vs. that of vector transfected control. ***p < 0.001 vs. MIPS-9541 control by Welch’s t-test.

Pharmacological evaluations of lead polymyxin analogues

We firstly assessed the MICs of the four lipopeptide analogues that have disrupted interaction with hPepT2. As shown in Table 2, FADDI-170 showed significantly reduced activity against FADDI-PA025, -EC006, -EC003, -EC001, -AB034 as well as A. baumannii ATCC 19606 and ATCC 17978 (MICs increased ≥16 folds, compared to polymyxin B). FADDI-175 displayed low activity against all the tested strains (MICs ≥32 µg/mL), which was not further considered for subsequent testing. FADDI-793 is less active than polymyxin B against A. baumannii ATCC 19606, ATCC 17978, FADDI-AB034, -EC006 and -EC003. FADDI-795 was the most active among them, showing comparative MICs against these bacterial strains except for FADDI-PA025, when compared to polymyxin B. Interestingly, FADDI-793 and FADDI-795 were more active than polymyxin B against K. pneumoniae ATCC 13883.

We then employed the mouse model established in our laboratory to examine the nephrotoxicity of the four antibacterial polymyxin analogues (16). Interestingly, FADDI-795 showed no observable nephrotoxicity; while the other three analogues possess mild kidney toxicity, comparable to polymyxin B (Table 2).

Overall, the transporter uptake study and pharmacological assessments of these polymyxin analogues provide critical mechanistic data on the interaction and transport of polymyxins by hPepT2. Importantly, our integrated computational and cell biology approach highlighted the feasibility to reduce polymyxin nephrotoxicity by disrupting their interactions with hPepT2. Moreover, hPepT2 is a viable target for designing novel polymyxin-like lipopeptide antibiotics with reduced nephrotoxicity.

Discussion

Life-threatening MDR bacterial infections rank among the world’s most pressing health threats (17), yet the discovery of antibiotics with novel mechanisms of action has stalled for decades. Polymyxins remain as a last-line therapy for MDR Gram-negative bacterial pathogens, particularly in low to middle income countries; however, their clinical use has been significantly limited due to nephrotoxicity (3, 4).

Our previous study demonstrated that polymyxins are substrates of hPepT2 which mediates the substantial reabsorption of polymyxins in renal proximal tubular cells, leading to nephrotoxicity (11). However, little is known about the interaction between polymyxin molecule and hPepT2. Notably, previous studies have reported that megalin mediates the cellular transport of polymyxins in the kidney via a receptor-mediated endocytosis mechanism (4, 18). Receptor-mediated endocytosis is closely linked to changes in signalling pathways; thus, disrupting the interaction between polymyxins and megalin may lead to unpredictable signalling consequences and complicate treatment outcomes. In contrast, transporter-mediated uptake does not involve such signalling changes, making manipulation of polymyxin-hPepT2 interactions a more favourable strategy for mitigating polymyxin-associated nephrotoxicity. Furthermore, our previous study demonstrated a direct structural interaction of polymyxins with the membrane of human kidney proximal tubular cells (19), showing that polymyxin B inserts its fatty acyl tail into the hydrophobic region of the cell membrane. The stereochemistry at position 3 and the hydrophobicity at positions 6 and 7 of polymyxin B were found to be critical for this interaction. These findings provide important insights into how polymyxins alter membrane dynamics and lipid organization in kidney cells, contributing to a better understanding of polymyxin-induced nephrotoxicity.

The current study developed the first SIR model of polymyxins with hPepT2, expediting the discovery of novel safer lipopeptide antibiotics. We firstly employed MD simulations to predict the binding of polymyxins to hPepT2 (Fig. 1 and 2) and identified several critical amino acid residues of hPepT2 in the recognition of polymyxin molecules. An outward-facing model of hPepT2 was firstly constructed, indicating the access of polymyxin molecules to the lateral opening gate of hPepT2 (Fig. 3). The distinct conformation of polymyxin B indicates that it situates in the central binding cavity of hPepT2. To further explore the conformational change of polymyxin molecules required for hPepT2 transport, all-atom MD simulations were conducted with varying the orientations of polymyxin B. Our results revealed that such a conformation involved a notable flip of polymyxin B enhances the interaction between its fatty acyl chain and the deeper regions of predicted transporter binding pocket of hPepT2. Notably, the subsequent transporter mutagenesis and molecular characterisations of hPepT2 validated our proposed SIR model. Our integrated computational and experimental findings provide the proof-of-concept that disrupting the interaction of polymyxin and hPepT2 is a novel strategy to reduce their nephrotoxicity.

As polymyxins are cationic at physiological pH, we strategically selected the negatively charged amino acids of hPepT2 (i.e. aspartic acid and glutamic acid) for further investigation. The significantly reduced uptake of polymyxin B by hPepT2 mutants E214A, D215A, D342A, E555A and E622A demonstrated the critical role of these negatively charged residues in transporting polymyxins (Fig. 5 and 6). Moreover, the kinetics results of these transporter mutants elucidated the mechanism underpinning of such functional impairment (Table 1). As both transporter cell surface expression and turn-over rate account for the maximum velocity Vmax of a transporter protein (10), it was critical to evaluate the protein expression of these transporter mutants in total cell and at cell surface (Fig. 7). Both the kinetic analysis and protein expression profiles (Table 3) showed different mechanisms of impaired polymyxin uptake by hPepT2. For example, it is highly likely that E214A reduced Vmax due to a reduced total cell and membrane protein expression, while D215A reduced polymyxin uptake mainly due to impaired binding affinity to polymyxins (i.e. increased Km). Noteworthy, the transporter turnover rate of D215A might increase to compensate for the reduced cell surface expression, thus its Vmax was overall increased. Similarly, the decreased polymyxin uptake mediated via D342A was primarily attributed to its reduced polymyxin binding affinity, while its decreased membrane protein expression and likely elevated transporter turnover rate resulted in an unchanged Vmax. In contrast, the dysfunction of E555A was mainly due to its reduced Vmax likely resulted from a reduced total cell and membrane protein expression, although its binding affinity to polymyxins is slightly increased (i.e. decreased Km). The kinetic parameters of E622A were unable to be determined due to its low polymyxin uptake and our data indicated that its low protein expression in total cell and at cell surface contributed to its impaired transport activity in polymyxin uptake. Notably, because some transporter mutants (e.g., D342A) exhibited both increased Km and reduced membrane expression, it is not possible to conclusively determine whether the decreased transporter function results from reduced binding affinity, lower protein expression, or a combination of both, which is a limitation of the current study. Nevertheless, our molecular assay results support the applicability of the bioinformatic predictions, as they allow us to focus the analysis on a limited set of transporter residues that are relevant to our SIR model.

To validate the MD predictions, we further investigated the transport function of these mutant transporters through their uptake of the classic hPepT2 substrate Gly-Sar (Fig. 5A). The uptake of Gly-Sar was dramatically decreased in the mutants of E214A, D215A, D317A, D342A and E622A, which is consistent with the reduced cell surface expression of E214A, D215A, D342A and E622A (Fig. 7). D317A is exceptional as it showed decreased uptake of Gly-Sar but maintained transport function in the case of polymyxin B, which indicated that the residue of D317 is only important for the uptake of Gly-Sar but not polymyxins. E555A is another exception, which remained functional for the uptake of Gly-Sar but reduced uptake of polymyxins. As E555A had a reduced total cell and cell surface expression, its binding affinity to Gly-Sar might increase to compensate for the impaired Vmax. Thus, the binding pocket of polymyxins does not seem to be identical to that of Gly-Sar. Therefore, the transport inhibition study was only employed as a positive control initially and not considered in the subsequent evaluation of the antibacterial activity of polymyxin analogues.

According to the predicted topology model of hPepT2 (10), E214A and D215A situate at the ECD between TM5 and TM6; D342A resides at the ECD between TM7 and TM8; E555A locates at the ECD between TM9 and TM10; and E622A is within TM10. It was reported that the large ECD between TM9 and TM10 interacts with TM1 to maintain transport function (20). In our study, we demonstrated that E555 residue mildly increased binding affinity to polymyxin B with a reduced Vmax, likely resulted from the decreased protein expression. These findings highlight the need of more in-depth mechanistic understanding of hPepT2-mediated peptide transport. Notably, E622A in TM10 has a very low transport function in the uptake of polymyxin B, which is largely due to its low protein expression in total cell and on the cell surface (Fig. 5 and 7). Therefore, this residue appears to be essential for preserving hPepT2 stability. Similar findings have been reported in other SLCs (21–23). According to the protein sequence alignment between hPepT2 and another key isoform of SLC15A subfamily, hPepT1, E595 of hPepT1 is the equivalent residue to E622 in hPepT2. It has been shown that the alternation of E595 residue to cysteine substantially impacted on transport function but not protein expression (24). This finding reflects the differing substrate specificities and transport capacities of hPepT1 and hPepT2 (10), highlighting the essential role of hPepT2 in mediating polymyxin transport.Upon validating the predictions of the SIR model of polymyxins and hPepT2, our study was advanced to employ a chemical biology approach to rationally design and in-house synthesize polymyxin analogues. With the hope to interrupt the interaction between hPepT2 and these analogues, we aimed to discover candidate lipopeptide antibiotics with reduced nephrotoxicity while similar or better antibacterial activity. Based on the SIR model, the Dab1, Dab3, Dab5, Dab8 and Dab9 residues of polymyxins are likely to be involved with its interaction with hPepT2 (Fig. 4). Thus, we synthesised FADDI-167, FADDI-170, FADDI-175, FADDI-793 and FADDI-795 with alanine substitutions on these Dab residues. We first examined the hPepT2-mediated uptake of these polymyxin analogues. Interestingly, we discovered that all but not FADDI-167 disrupted the uptake via hPepT2 (Fig. 8). This finding indicated that FADDI-170, FADDI-175, FADDI-793 and FADDI-795 possibly impaired their uptake via hPepT2, thereby leading to reduced accumulation in the kidney and nephrotoxicity. Thus, these four analogues were proceeded for further pharmacological evaluations. Interestingly, FADDI-795 and FADDI-175 showed the best and least antibacterial activity among them, respectively (Table 4). Finally, we employed our mouse model to evaluate the nephrotoxicity of these lipopeptides. Using histological evaluations, FADDI-795 showed no observable toxicity in mouse kidneys and stands out as the lead compound with comparable antibacterial activity. Although beyond this study, FADDI-795 is subject to further pharmacological evaluations, in particular pharmacokinetics in our drug discovery program.

Overall, our study developed the first SIR model of polymyxin with hPepT2 by integrating computational, chemical biology and molecular approaches. Our molecular studies provide strong experimental support for the bioinformatic predictions. Although D215 was the only residue experimentally confirmed to play a key role in polymyxin binding, this insight enabled the targeted design of polymyxin analogues. Subsequent studies of drug efficacy and toxicity demonstrated that FADDI-795 retains antibacterial activity and, importantly, shows reduced nephrotoxicity, highlighting the potential of our approach and underscoring the significance of these findings for drug discovery. Our experimental findings validated the first SIR model and enhanced our understanding of the molecular interaction between the polymyxin molecule and hPepT2. These mechanistic insights are essential for elucidating the clinical dose-limiting nephrotoxicity associated with polymyxin treatment. Moreover, this work serves as a proof-of-concept study, providing the first experimental validation of the predicted polymyxin-hPepT2 binding model. Our ongoing drug discovery efforts will build on these findings to refine the predictive model and accelerate lead optimization for the development of novel, safer lipopeptide antibiotics against Gram-negative bacterial infections.

Materials and Methods

Reagents and chemicals

[3H] glycylsarcosine (Gly-Sar; 2 Ci/µmol) was obtained from BioScientific Pty. Ltd. (Gymea, NSW, Australia). Dulbecco’s Modified Eagle Medium (DMEM), fetal bovine serum (FBS), and other culture supplements were purchased from Thermo Scientific (Lidcombe, NSW, Australia). Gly-Sar and polymyxin B1 for molecular studies were bought from Sapphire Biosciences (Redfern, NSW, Australia). Polymyxin B sulfate for animal studies was purchased from BetaPharm (Shanghai, China). MIPS-9541 was synthesised as previously reported (11, 25, 26). Plasmids containing the full-length hPEPT2 and hPepT2-c-Flag were obtained from Australian Biosearch (Balcatta, WA, Australia). All other reagents and chemicals were purchased from Sigma-Aldrich (Castle Hill, NSW, Australia).

Coarse-grained molecular dynamics (MD) simulations

Polymyxin B1 was employed as the prototype (27) and the coarse-grained (CG) model system of hPepT2 was constructed using the Martini Maker module in CHARMM-GUI (28). The membrane lipid composition was 50% phosphatidylcholine, 25% phosphatidylethanolamine and 25% phosphatidylserine (29). In each simulation system, 10 coarse-grained polymyxin molecules were randomly introduced into the simulation system to enhance the binding probability of polymyxin B to hPepT2. These were solvated using CG water particles and neutralised with 0.1 M sodium chloride in a 15 × 15 × 14 nm3 simulation box. All CG simulations were conducted under periodic boundary conditions using the GROMACS program (version 2021.2) and Martini 2.2 force field (30, 31). To eliminate steric interference, the steepest energy minimization was performed for every system to give the maximum force below 1,000 kJ mol−1 nm−2. Subsequently, a six-step equilibration cycle was carried out by gradually turning off the position restraints on the lipid and polymyxin molecules. Finally, 3μs-production simulations with two replicates were conducted for each system in the NPT (referring to the number of particles (N), system pressure (P) and temperature (T)) ensemble. The temperature and pressure were coupled to 310 K using the velocity rescaling method (time constant of 1 ps) and 1 bar using the semi-isotropic barostat and Parrinello-Rahman algorithm (time constant of 5 ps), respectively (32).

All-atom MD simulations

The previous study of PepTXc, a homolog of hPepT2, revealed that the protonation of its His67 promotes the outward opening conformation (33), therefore the corresponding residue His87 of hPepT2 was set to be protonated during all-atom simulations to bias toward an outward opening state. In the two CGMD replicates, polymyxin B eventually bound to the lateral opening gate of hPepT2. To examine the detailed interaction between polymyxin and hPepT2, a single polymyxin B molecule was introduced to the binding pocket of hPepT2 using Autodock Vina (34). Furthermore, to enhance the sampling of the binding dynamics, four different poses of polymyxin B were employed to build the initial hPepT2-polymyxin complex. Then, the membrane protein system of hPepT2-polymyxin was constructed using the membrane builder module in CHARMM-GUI (35).

GROMACS 2021.2 was used to perform all MD simulations with the CHARMM36 all-atom force field (36). A time step of 2 fs was employed to integrate the Newtonian equations of motion and all bonds involving hydrogen atoms were constrained. Long-range electrostatic interactions were treated using the Particle Mesh Ewald (PME) method with a switching function applied between 10 and 12 Å (37). After energy minimization, the system was equilibrated through a series of MD simulation steps in the NVT (referring to the number of particles (N), system volume (V) and temperature (T)) and NPT ensembles (310 K, 1 bar), during which positional restraints on membrane and protein atoms were gradually reduced. The root mean square deviations of polymyxin B were calculated to examine the simulation equilibration (Fig. S1). No restraints were applied in the final stage of equilibration and throughout the production run. The PME method was employed to treat long-range electrostatic interactions with a short-range cut-off of 1.2 nm, while the shifted Lennard–Jones potential algorithm was used to calculate Van der Waals interactions with a general cut-off of 1.2 nm and a shifting cut-off of 1.0 nm (38, 39). The trajectory during production simulations was recorded every 10 ps.

Site-directed mutagenesis and plasmid DNA sequencing

The hPepT2 mutants were generated using a QuickChange II site-directed mutagenesis kit (Agilent Technologies, Mulgrave, Victoria, Australia) with the primers listed in Table 4. All mutant construct sequences were confirmed by DNA sequencing (Ramaciotti Centre for Gene Function Analysis, Randwick, NSW, Australia) with the BigdyeTM Terminator v3.1 kit (Thermo Scientific, Lidcombe, NSW, Australia) (9, 21, 40–42).

Transfection of hPepT2 and its mutants into HEK293 cells

HEK293 cells were cultured in DMEM supplemented with 10% FBS at 37°C with 5% CO2. Cells were transfected with hPepT2 or its mutant plasmid DNA using Lipofectamine 2000 Reagent as to the manufacture’s instruction (Invitrogen, Mount Waverley, VIC, Australia) (7, 11, 43–45). At 24 h after transfection, the cellular uptake of [3H] Gly-Sar or MIPS-9541 was assessed.

Transport uptake assay

To evaluate the uptake of [3H] Gly-Sar via hPepT2 or its mutant transporters, we measured the accumulation of [3H] Gly-Sar in over-expressing HEK293 cells. The uptake was conducted at 37°C in phosphate-buffered saline (PBS, pH 5.0) containing 5 mM glucose. The specific substrate concentration was 2.5 μM [3H] Gly-Sar and the duration was 8 min according to our previous studies (7, 11). The uptake assay was terminated by three rapid washes with ice-cold PBS. Samples were lysed in 0.2 M NaOH, neutralized in 0.2 M HCL and then processed for liquid scintillation counting. The uptake counts for hPepT2 or its mutant expressing cells were all subtracted with that of the control.

To evaluate the uptake of polymyxin B via hPepT2 or its mutant transporters, we measured the accumulation of a validated fluorescent polymyxin probe MIPS-9541 and polymyxin analogues (i.e. FADDI-167, -170, -175, -793 and -795, Fig. S2) in over-expressing HEK293 cells (11). The uptake was performed in PBS (pH 5.0) supplemented with 5 mM glucose for 10 min at 37°C. Intracellular fluorescence accumulation was quantified with excitation/emission wavelengths of 350 nm / 518 nm using a Tecan Safire II microplate reader (Thermo Scientific, Lidcombe, NSW, Australia). Background counts of the vector-transfected cells were subtracted from all uptake measurements. Kinetic analysis was conducted by measuring the uptake of transporter expressing cells with a range of concentrations of MIPS-9541 (0 - 50 µM) over 10 min. Apparent Km and Vmax values were determined using GraphPad Prism 10.0 (11, 46).

Fluorescence imaging of MIPS-9541

HEK293 cells were seeded on chamber slides and transfected with hPepT2 or its mutant plasmids. After 24 h, culture medium was removed and cells were incubated with freshly prepared MIPS-9541 solution in PBS (10 µM, pH 5.0) for 10 min at 37°C. Cells were then washed with ice-cold PBS for three times and mounted in SlowFade Gold Antifade Mountant reagent supplemented with 4’,6-diamidino-2-phenylindole (DAPI, Thermo Scientific, Lidcombe, NSW, Australia). Samples were imaged with a Leica Thunder 3D Imager (Leica Microsystems, North Ryde, NSW, Australia).

Cell surface biotinylation

We employed sulfo-NHS-SS-Biotin to label the plasma membrane protein in HEK293 cells overexpressing hPepT2 or its mutants (47). Cell culture plates were pre-chilled on ice. Culture medium was aspirated, and cells were washed twice with cold PBS (pH 8.0). Freshly prepared 1 mg/mL Sulfo-NHS-SS-Biotin in PBS (pH 8.0) was then added to each well and incubated for 30 min. Cells were washed twice with 100 mM glycine in PBS (pH 8.0) and then three times with PBS (pH 7.4). Cells were lysed in the lysis buffer (Tris 10 mM, NaCl 150 mM, EDTA 1 mM, SDS 0.1% and Triton X-100 1% with 1:1,000 dilution of a protease inhibitor cocktail) (21). Cell lysate was subject to centrifugation at 14,000 g and 4°C for 10 min and the supernatant was collected. Pre-washed Streptavidin-Agarose beads were supplied to the samples with rotating at 4°C for 1 h. After three rounds of centrifugation and washing with PBS (pH 7.4), beads were resuspended in PBS with 1x Laemmli buffer containing 2.5% β-mercaptoethanol and incubated at 55°C for 30 min. The remaining supernatant was collected, denatured and subject to electrophoresis.

Total cell lysis, electrophoresis and immunoblotting

Upon washing, cells were collected and lysed in the lysis buffer. After incubating on ice for 10 min, cell lysate was centrifugated at 14,000 g and 4°C for 10 min. The supernatant was collected as total cell lysate and denatured at 55°C in 1x Laemmli buffer containing 2.5% β-mercaptoethanol for 30 min.

Protein lysate samples were loaded to 10% SDS-PAGE gels followed with electrophoresis as described before (48) and the protein samples on the gels were transferred to polyvinylidene difluoride (PVDF) membranes. Following blocking with 5% non-fat milk in PBS-Tween (Na2HPO4 80 mM, KH2PO4 20 mM, NaCl 100 mM, and 0.05% Tween 20, pH 7.5) and multiple washing with PBS-Tween, the immunoblots were incubated at 4°C overnight with anti-flag (DYKDDDDK) antibody (1:1,000; Catalogue Number 2368, Genesearch, Arundel, QLD, Australia) or anti-β-actin antibody (1:2,000, Catalogue Number 4967, Genesearch). After washing with PBS-Tween for three times, the immunoblots were incubated at room temperature for 1 h with goat anti-rabbit IgG conjugated with HRP (1:5,000; Catalogue Number A0545, Sigma-Aldrich) or goat anti-mouse IgG conjugated with HRP (1:10,000; Catalogue Number A2554, Sigma Aldrich). The blots were washed, incubated with Immobilon Western Chemiluminescent HRP substrate (Merck Millipore, Kilsyth, VIC, Australia) and imaged with an ImageQuant LAS 500 (GE healthcare, Avantor, PA, USA). The densitometry analysis of immunoblots was processed with Image J (NIH, Bethesda, MD, USA).

Chemical synthesis of polymyxin analogues

Five lipopeptides (FADDI-167, FADDI-170, FADDI-175, FADDI-793 and FADDI-795) were synthesised and purified in house as described before (16, 49). Briefly, the lipopeptides were prepared using a Protein Technologies Prelude automated peptide synthesizer as to the standard protocol of Fmoc solid-phase peptide chemistry.

Measurements of minimum inhibitory concentrations (MICs)

MICs were measured using the broth micro-dilution method with clinical isolates and American Type Culture Collection (ATCC) strains (50). Each bacterial strain (∼106 colony forming units per mL) was suspended in 100 μL of freshly prepared cation-adjusted Mueller-Hinton broth (CaMHB) in 96-well plates. Lipopeptides were diluted into different concentrations in CaMHB and added to the plates for subsequent incubation at 37°C for 18-24 h. MICs were determined as the lowest concentration that the visible growth of bacteria was inhibited.

Nephrotoxicity in mice

The mouse nephrotoxicity study was approved by the Monash Animal Ethics Committee. Mice were maintained in micro-isolators in a temperature-controlled PC2 animal laboratory with an ambient humidity of 50-70% and a 12 h/12 h dark/light cycle. Lipopeptide solutions were subcutaneously administered to mice with a dose of 8 mg/kg every 2 h for 6 doses in one day (a total dose of 48 mg/kg/day). At 24 h after the last injection, mice were sacrificed, the kidneys were collected from mice immediately and fixed in 10% formalin buffer (pH 7.4, Sigma Aldrich, Australia). Histological examination of the kidney samples was conducted at the Australian Phenomics Network-Histopathology and Organ Pathology (University of Melbourne, Parkville, VIC, Australia). An experienced pathologist, who was blind to the treatment, rated the histology and severity of the nephrotoxicity as to three grades with scores 1 to 10 (16). The percentage of kidney affected was also scaled in the range of 0-6 (0: <1%; 1: 1 to<5%; 2: 5 to <10%; 3: 10 to <20%; 4: 20 to<30%; 5: 30 to ≤ 40% and 6: >40%). A semi-quantitative score (SQS) was calculated using the grade score and kidney damage percentage to quantify the overall kidney damage (16).

Statistical analyses

Unpaired t-test was employed to detect the variance between the protein expression of hPepT2 and its mutants. Welch’s t-test was used to assess the differences between the transport function or kinetic parameters of hPepT2 and its mutants. All data are expressed as mean ± standard deviation (SD) with a p value of <0.05 considered as statistically significant. All experiments were repeated three times with triplicates in each experiment.

eLife Assessment

This study presents a potentially important structure-interaction model describing recognition of polymyxin antibiotics by the human transporter hPepT2 and applies these insights to guide the rational design of potent polymyxin B analogues with reduced nephrotoxicity. The conclusions are supported by compelling evidence, integrating molecular dynamics simulations, transporter mutagenesis, uptake assays, protein expression analyses, antibacterial testing, and mouse nephrotoxicity studies, although some mechanistic interpretations remain limited by altered transporter expression and the absence of direct structural validation.

Reviewer #1 (Public review):

Summary:

Polymyxins are the last line of drugs to treat gram-negative bacteria-induced multi-drug resistance; however, they cause nephrotoxicity in 60% of patients. In this work, the authors have studied the structure-interaction relationship (SIR) of polymyxins with hPepT2 using computational and experimental methods. Moreover, it is observed that the electrostatic interactions coordinate the hPepT2-Polymyxin interactions; hence, an alanine scanning strategy is used to understand the interactions and derive the polymyxin variants.

Computational methods such as molecular modeling, coarse-grained and all-atom MD simulations, and interaction studies are performed, while the results are validated in the mouse model, which is a great strategy to prove the hypothesis.

Strengths:

A clear understanding of the hPepT2-Polymyxin interactions and the role of electrostatic interactions is one of the very important strengths of the paper. In addition, this work proposes a great pipeline for using computational approaches and experimental validation methods to guide the development of newer antibiotics.

Overall, the study proposes novel polymyxin analogues with reduced or no nephrotoxicity, thereby providing a promising foundation for the rational development of safer lipopeptide antibiotics.

Weaknesses:

This work is very well executed and presented; however, addressing the following concerns might improve the presentation of the work:

(1) The introduction is well articulated; however, including a paragraph on the known inhibitors might be helpful in understanding the current status. In addition, it might also help to introduce Dabs, FADDI variants, Gly-sar and MIPS.

(2) The following details of modeling with AlphaFold2 should be included: how the final structure was selected, what the RMSD and structure alignment of the template are, and the final selected structure. A section on modeling with all the parameter details might be useful for reproducing the structure. In addition, specify how the alanine scanning was performed alongside the structure prediction of polymyxins.

(3) In the all-atom MD simulation method, detailing several parameters might help in reproducing the results: simulation time for each system, water model, system composition, protonation state, box type and dimensions, salt ions and concentration, membrane parameters and ligand parameterization methods. Also, the following details on energy minimization might be useful: minimization algorithm, number of steps for minimization and structure restraints in place.

(4) On page 6, line 210, the MIC is used for the first time; although MIC is given in the abbreviation list, the first occurrence should have a complete name. A one-line explanation of MIC in the introduction or wherever suitable might be better but is not mandatory.

(5) Similarly, Gly-sar is first mentioned on page 8, line 301, but its complete name is only mentioned later on page 10, line 368. This can be addressed if a short description is included in the introduction section.

(6) For coarse-grained MD simulation, why were 2 replicates performed? Most studies perform 3 replicates, which are also good in terms of statistics and error bar calculations. In addition, the authors should specify whether an independent minimization is done for each of the two replicates or whether the minimization step is common for both.

(7) For MD simulation results, giving simulation movies in supplementary results might be a better way to show how the trajectories behaved.

(8) The description of visualisation software such as VMD or PyMol is missing. The authors should specify if any visualization tool is used.

(9) For the mouse model study, the authors claim that FADDI-795 has no observable nephrotoxicity; however, the n=3 shows that a very small number of mouse models were used to make the assumption. In addition, the number of mice used in each experiment is not explicitly mentioned in the methods section.

(10) In Table 2, the column 8 header is not visible.

Reviewer #2 (Public review):

Summary:

Jiang et al. sought to elucidate the molecular basis of polymyxin antibiotic interaction with the renal transporter hPepT2, a transporter previously implicated in polymyxin-induced nephrotoxicity. They combined molecular dynamics simulations with transporter mutagenesis, functional uptake assays, kinetic analyses, protein expression studies, antibacterial susceptibility testing, and mouse nephrotoxicity experiments to develop a structure-interaction relationship (SIR) model and apply this model to the rational design of polymyxin analogues.

Overall, the study represents a substantial multidisciplinary effort that integrates computational and experimental approaches. The identification of transporter residues involved in polymyxin recognition and the subsequent design of analogues with reduced hPepT2-mediated uptake provide a valuable framework for developing safer polymyxin antibiotics. In particular, the identification of FADDI-795 as an analogue that retains antibacterial activity while exhibiting reduced nephrotoxicity represents an encouraging proof of concept.

Strengths:

The computational predictions are strengthened by extensive experimental validation, including site-directed mutagenesis, transport kinetics, fluorescence uptake assays, membrane expression analyses, and in vivo toxicity studies. The consistency between multiple independent experimental approaches increases confidence in many of the authors' conclusions.

Weaknesses:

Several conclusions would benefit from a more cautious interpretation. A major limitation is that several transporter mutations substantially altered total or membrane protein expression, making it difficult to distinguish effects on substrate binding from indirect effects caused by impaired transporter stability or trafficking. The authors acknowledge this limitation in the Discussion, but some mechanistic conclusions remain stronger than the available evidence supports.

Similarly, while the proposed binding model is biologically plausible and supported by mutagenesis, it remains an inferred model derived from molecular simulations rather than a direct structural determination. Statements describing the model as "validated" should therefore be moderated to indicate that the experimental data provide support rather than definitive structural confirmation.

The translational implications are promising but remain preliminary. Although FADDI-795 demonstrated reduced nephrotoxicity in the mouse model while maintaining antibacterial activity, no pharmacokinetic studies were presented to demonstrate reduced renal accumulation or altered tissue distribution, and additional efficacy studies in infection models would further strengthen the therapeutic claims.

Reviewer #3 (Public review):

Summary:

Jiang et al. described findings aimed at interrogating the interactions of the antibiotic polymyxin B with human kidney proteins that mediate nephrotoxicity. Their findings using both computational molecular dynamics simulations and experimental approaches illustrate the importance of aspartic acid residues (D215) in mediating the antibiotic uptake into the cells, and upon mutagenesis with Alanine, the effects are less pronounced. Further, they could modify the antibiotic units interacting with proteins into less toxic peptides with retained antibacterial properties.

Strengths:

I was impressed by this text, which advances the knowledge of how the antibiotic causes human nephrotoxicity and how this could be exploited into less problematic antibiotic peptides.

Weaknesses:

Interactions of Polymyxin B with kidney proteins were not demonstrable in vivo, and with reliable technologies such as X-ray or NMR.

Author response:

The following is the authors’ response to the original reviews.

Reviewer #1 (Public review):

(1) The introduction is well articulated; however, including a paragraph on the known inhibitors might be helpful in understanding the current status. In addition, it might also help to introduce Dabs, FADDI variants, Gly-sar and MIPS.

We thank the reviewer for the suggestion. We have included a paragraph on the substrates of hPepT2 (Lines 104-112).

(2) The following details of modeling with AlphaFold2 should be included: how the final structure was selected, what the RMSD and structure alignment of the template are, and the final selected structure. A section on modeling with all the parameter details might be useful for reproducing the structure. In addition, specify how the alanine scanning was performed alongside the structure prediction of polymyxins.

We appreciate the reviewer's careful assessment of our computational methodology. In the Materials and Methods section of the revised manuscript, we have added a dedicated subsection describing the detailed methods: "Structural modelling of hPepT2 and polymyxins" as shown below (Lines 435-447).

(1) hPepT2 structure prediction: The inward-open conformation of hPepT2 was predicted using AlphaFold2 via the ColabFold implementation with default parameters, employing the MMseqs2 multiple sequence alignment pipeline. To obtain the physiologically relevant outward-open conformation of hPepT2 for substrate binding, homology modeling was performed using MODELLER with the cryo-EM structure of rabbit PepT2 in the outward-open state (PDB: 7NQK) as the template. The sequence alignment between hPepT2 and rabbit PepT2 was conducted using Clustal Omega. One hundred models were generated, and the final model was selected based on the lowest Discrete Optimized Protein Energy score and verified by Ramachandran plot analysis (with >95% of residues in favored regions).

(2) Polymyxin structure and alanine scanning: The structure of polymyxin B1 was constructed and energy-minimized using the CHARMM36 force field. Computational alanine scanning of polymyxin B1 was performed by individually replacing each Dab (2,4-diaminobutyric acid) residue at positions 1, 3, 5, 8, and 9 with alanine using the mutagenesis wizard in PyMOL.

(3) In the all-atom MD simulation method, detailing several parameters might help in reproducing the results: simulation time for each system, water model, system composition, protonation state, box type and dimensions, salt ions and concentration, membrane parameters and ligand parameterization methods. Also, the following details on energy minimization might be useful: minimization algorithm, number of steps for minimization and structure restraints in place.

We thank the reviewer for the suggestion. In this study, all-atom molecular dynamics simulations were performed using the TIP3P water model for solvation. The system was placed in a rectangular simulation box with dimensions of 11 × 11 × 14 nm3. Neutralization was achieved by adding 0.15 M NaCl, resulting in a total of approximately 158,000 atoms. The membrane lipid composition consisted of 50% phosphatidylcholine, 25% phosphatidylethanolamine, and 25% phosphatidylserine, consistent with the coarse-grained molecular dynamics simulations. Force field parameters for polymyxin B1 were generated using CGenFF. Energy minimization was carried out using the steepest descent algorithm with default parameters. For production simulations, 500-ns simulations were run for sampling the outward‑open conformation, and 100-ns simulations were run for analyzing the polymyxin‑hPepT2 interaction. We have provided these methodological details in the revised manuscript (Lines 477-492).

(4) On page 6, line 210, the MIC is used for the first time; although MIC is given in the abbreviation list, the first occurrence should have a complete name. A one-line explanation of MIC in the introduction or wherever suitable might be better but is not mandatory.

We thank the reviewer for the suggestion. We have included the full term of MIC and its definition in the revised manuscript (Lines 236-238).

(5) Similarly, Gly-sar is first mentioned on page 8, line 301, but its complete name is only mentioned later on page 10, line 368. This can be addressed if a short description is included in the introduction section.

We apologize for the overlook. We have now referred to the full term of Gly-Sar and a short description on Page 3 (Lines 104-106), when it was first mentioned.

(6) For coarse-grained MD simulation, why were 2 replicates performed? Most studies perform 3 replicates, which are also good in terms of statistics and error bar calculations. In addition, the authors should specify whether an independent minimization is done for each of the two replicates or whether the minimization step is common for both.

In our study, coarse-grained MD simulations served as a preliminary exploration to identify the potential binding trajectory of polymyxin B to hPepT2, rather than the primary source of quantitative interaction data. The two independent CG-MD replicates yielded highly consistent final binding poses of polymyxin B at the lateral opening gate of hPepT2, serving this exploratory purpose. Furthermore, more detailed interaction energetics were subsequently characterized using all-atom MD simulations with four replicates (Fig. 4). We have now clarified in the revised Methods that independent energy minimization was performed for each of the two CG-MD replicates (Lines 465-469).

(7) For MD simulation results, giving simulation movies in supplementary results might be a better way to show how the trajectories behaved.

We thank the reviewer for the suggestion. We have prepared an animation of the coarse-grained MD trajectory showing the binding trajectory of polymyxin B to hPepT2, which has been submitted as a Supplementary Movie (.mp4 format). This movie visualizes how polymyxin B molecules spontaneously approach and bind to the lateral opening gate of hPepT2 over the 3-μs simulation. The Supplementary Movie is referenced in the revised Methods section (Lines 471-474).

(8) The description of visualisation software such as VMD or PyMol is missing. The authors should specify if any visualization tool is used.

We apologize for this omission. In the revised manuscript, we have specified in the Methods section that molecular visualizations, structural figures, and trajectory animations were prepared using PyMOL (Schrödinger) and ChimeraX (Lines 471-474).

(9) For the mouse model study, the authors claim that FADDI-795 has no observable nephrotoxicity; however, the n=3 shows that a very small number of mouse models were used to make the assumption. In addition, the number of mice used in each experiment is not explicitly mentioned in the methods section.

We have now added “n=3 each group” in the Methods section (Line 603). In this proof-of-concept study, we employed acute kidney damage analysis in a small number of mice to rapidly screen these analogues for nephrotoxicity. Three biological replicates were used because our mouse nephrotoxicity model is robust as demonstrated by a large number of animals in our drug discovery program (e.g. Nature Communications 2022, 13, 1625). Moreover, minimising animal use is a fundamental principle of the 3Rs (Replacement, Reduction, and Refinement).

(10) In Table 2, the column 8 header is not visible.

We apologize for the format error and have now fixed it in Table 2.

Reviewer #1 (Recommendations for the authors):

(1) The MD simulation trajectories or movies are not provided.

As mentioned in our response to Comment 7 above, a Supplementary Movie (mp4 format) showing the CG-MD binding pathway has been provided with the revised manuscript and referenced in the Methods section.

(2) Mouse model experiments are performed over n=3, which might not be statistically significant in case of an experiment with a large n.

Please see our response to Comment 9 above.

Reviewer #2 (Public review):

(1) Several conclusions would benefit from a more cautious interpretation. A major limitation is that several transporter mutations substantially altered total or membrane protein expression, making it difficult to distinguish effects on substrate binding from indirect effects caused by impaired transporter stability or trafficking. The authors acknowledge this limitation in the Discussion, but some mechanistic conclusions remain stronger than the available evidence supports.

We agree with the reviewer that it is difficult to draw definite conclusions when both Km and Vmax are altered in some mutants, especially when expression of the mutant transporter is impaired. However, we consider kinetic analysis to be more indicative of altered substrate binding, as changes in Km are generally more closely associated with substrate recognition and binding, whereas reduced transporter expression predominantly affects Vmax. Nonetheless, to avoid overinterpretation, we have moderated our conclusions by replacing “was/is” with “may be” (such as Lines 318, 319, 323, and 347).

(2) Similarly, while the proposed binding model is biologically plausible and supported by mutagenesis, it remains an inferred model derived from molecular simulations rather than a direct structural determination. Statements describing the model as "validated" should therefore be moderated to indicate that the experimental data provide support rather than definitive structural confirmation.

We thank the reviewer for the suggestion and have replaced “validated” with “explored” in the main text (Lines 298).

(3) The translational implications are promising but remain preliminary. Although FADDI-795 demonstrated reduced nephrotoxicity in the mouse model while maintaining antibacterial activity, no pharmacokinetic studies were presented to demonstrate reduced renal accumulation or altered tissue distribution, and additional efficacy studies in infection models would further strengthen the therapeutic claims.

We thank the reviewer for the suggestion. The current study is a proof-of-concept investigation to determine whether the structure-interaction relationship (SIR) model of hPepT2 and polymyxin can be used to design new lipopeptides with reduced nephrotoxicity. Following screening of the newly designed lipopeptides, we will select the most promising candidate for comprehensive pharmacological evaluations, including pharmacokinetics, toxicology, and efficacy.

Reviewer #2 (Recommendations for the authors):

(1) Strengthen the interpretation of the mutagenesis data. For transporter mutants that exhibit reduced total or cell surface expression, consider discussing more explicitly the extent to which reduced polymyxin uptake may result from altered protein stability or trafficking rather than direct effects on substrate recognition. Where possible, clarify this distinction throughout the Results and Discussion.

We thank the reviewer for the recommendation. As suggested, we have expanded the discussion on the mechanisms underlying the altered protein expression of the hPepT2 mutants (Lines 325-333).

(2) Moderate statements regarding structural validation. The molecular dynamics simulations, mutagenesis, and functional studies provide strong support for the proposed structure-interaction model; however, wording such as "validated the first SIR model" could be softened to reflect that the binding model remains computationally inferred rather than directly resolved by structural methods.

We thank the reviewer for the recommendation. We have revised the statement as described in our response to Comment 2 above.

(3) Expand the discussion of study limitations. A more explicit discussion of the limitations associated with molecular dynamics predictions, the influence of altered transporter expression on functional interpretation, and the lack of direct pharmacokinetic measurements of renal accumulation would improve the balance of the manuscript.

We thank the reviewer for the recommendation. We have expanded the discussion by elaborating the limitations of the molecular dynamics predictions (Lines 289-291), the effect of altered transporter expression on functional interpretation (Lines 324-333), and the lack of direct pharmacokinetic and renal accumulation measurements (Lines 393-401).

(4) Provide additional methodological detail where appropriate. Please clarify the number of biological replicates used for each experimental approach, report exact P values where practical, and indicate whether assumptions for the statistical tests were assessed.

We thank the reviewer for the recommendation. We have provided more information on experimental replicates in the figure legends. P values have been added (Lines 208-217), and the statistical tests used for every analysis are now specified in the corresponding table or figure legends.

(5) Future validation of the lead compound. Although this may fall outside the scope of the present manuscript, it would be helpful to briefly discuss future studies aimed at evaluating the pharmacokinetics, renal exposure, and efficacy of FADDI-795 in relevant infection models, as these will be important for establishing its translational potential.

We thank the reviewer for the recommendation. We have now incorporated this discussion into the revised manuscript (Lines 396-399).

(6) Improve figure presentation. Several figure legends could include additional methodological information to make the figures more self-contained. In particular, indicating sample sizes, statistical tests used, and definitions of error bars would improve readability.

We thank the reviewer for the recommendation. We have added this information to the figure legends as suggested.

(7) Language and style. The manuscript would benefit from careful language editing to improve grammar, sentence structure, and readability. Several long sentences in the Discussion could be shortened, and minor typographical errors should be corrected throughout the manuscript.

We thank the reviewer for the recommendation. We have revised the long sentences where possible and carefully proofread the manuscript to improve grammar, sentence structure, and readability. The revised manuscript has also been reviewed by a native English speaker.

(8) Terminology. Please ensure that abbreviations such as "SIR" are defined at first use and are used consistently throughout the manuscript.

We have confirmed that SIR is defined when first introduced in both the Abstract and main text (Lines 47 and 120).

(9) Minor corrections.

- Check for consistency in the naming of hPepT2/hPEPT2 throughout the manuscript.

- Verify that all figures, supplementary figures, and tables are cited sequentially in the text.

- Consider reporting confidence intervals alongside kinetic parameters (Km and Vmax) where appropriate to facilitate interpretation.

We thank the reviewer for the recommendation. We have revised the manuscript accordingly. “hPepT2” is now used consistently throughout the text. We have revised the order of the supplementary figures to ensure that they are cited sequentially in the text; and included the 95% confidence interval of the Km and Vmax values of the D215A mutant compared to the wild type (Lines 208-212).

Reviewer #3 (Public review):

(1) Interactions of Polymyxin B with kidney proteins were not demonstrable in vivo, and with reliable technologies such as X-ray or NMR.

We thank the reviewer for the comment. To the best of our knowledge, polymyxin-induced nephrotoxicity is primarily driven by renal tubular accumulation, followed by mitochondrial dysfunction, oxidative stress, inflammatory activation, and apoptosis, ultimately resulting in proximal tubular injury. Little is known about the molecular interactions of polymyxins with kidney proteins.

Reviewer #3 (Recommendations for the authors):

(1) Title: I suggest that you bring out the actual findings to convey the message.

We thank the reviewer for the comment. We have changed the title to “Lateral opening site of human oligopeptide transporter 2 plays a key role in the interaction with polymyxins”.

(2) Line 133: Provide data for these 10 molecules (possibly by supplementary file/table).

We thank the reviewer for the suggestion. As described in the Methods, ten coarse-grained polymyxin molecules were randomly inserted into the upper water layer of the simulation system (Lines 452-456). Because these molecules are structurally identical and share the same force field parameters, they differ only in their initial random positions. Thus, we believe that presenting the coordinates or labelling the individual molecules would not provide additional information and can be misleading.

(3) Line 135: Figure S3 mentioned prior Figure S2.

We apologize for the error. We have now revised the sequence of the supplementary figures.

(4) Lines 159 - 171: Cite Figure 5 a - b appropriately. Describe the obtained results fully.

We thank the reviewer for the suggestion. We have expanded the relevant section to include a more detailed description of the results, with reference to Fig. 5a and 5b (Lines 187-193).

(5) Line 208: Was cytotoxicity performed to validate this claim?

We did not specifically evaluate the cytotoxicity of the lipopeptides in cultured cells because the cellular uptake data already indicated their potential toxicity. Importantly, histopathological assessment in animal models provides the most reliable evaluation of their in vivo toxicity (Nature Communications 2022, 13, 1625).

(6) Line 509: Provide the approval number.

The animal ethics approval number is AEC37419, which has been added to the “Ethics approval and consent to participate” section.

(7) Table 2: Add the Standard deviation and the number of replicates performed.

We thank the reviewer for the suggestion. We have added “n = 3 mice per group” to the table legend. As described in the Methods (Lines 590-597), the MIC of each lipopeptide was determined. MIC values are reported as absolute concentrations rather than continuous variables; therefore, reporting the mean ± standard deviation is not applicable. Likewise, the maximum kidney Semi-Quantitative Score (SQS) was determined by histological examination and is presented as the representative maximum score for each treatment group. Accordingly, the calculation of a mean and standard deviation is not appropriate for either MIC or SQS values presented in Table 2.

(8) The authors need to indicate the number of replicates performed in the methods section.

We thank the reviewer for the suggestion. As mentioned in our response to Recommendation 6 from Reviewer 2 above, we have added this information to the figure legends as suggested.

Version 2

Abstract

Background

Multidrug-resistant (MDR) Gram-negative bacteria have presented a critical global health crisis and polymyxins remain a last-line therapy. However, their clinical use has been largely limited by nephrotoxicity. Human oligopeptide transporter 2 (hPepT2) is a membrane transporter mediating the reabsorption of polymyxins in renal cells, contributing to their nephrotoxicity. However, it remains unclear how polymyxins interact with hPepT2.

Methods

We investigated the structure-interaction relationship (SIR) of polymyxins with hPepT2 by integrating computational, chemical, and cell biology approaches. Bioinformatic modelling predicted the essential residues of hPepT2 for the binding to polymyxins. Transporter mutagenesis and molecular analysis were employed to examine the role of each residue in the interaction between hPepT2 and polymyxins. Subsequently, we synthesised a series of new analogues with alterations to the moieties of the polymyxin scaffold that are critical for binding with hPepT2, and assessed their antibacterial activity and nephrotoxicity.

Results

Our bioinformatic modelling proposed an outward-facing structure of hPepT2 with a possible transport pathway, in which polymyxins bind to the lateral opening site of hPepT2, particularly E214, D215, D317, D342, and E622. Molecular assays for transporter function and expression confirmed that D215 of hPepT2 is critical for polymyxin binding, while several other residues significantly impact transporter turnover rate and/or protein expression. Our experimental validations showed that the lipopeptides with alterations to the Dab1, Dab3, Dab5, and Dab9 residues of polymyxins demonstrated decreased interactions with hPepT2. Among these analogues, alanine substitution at Dab3 showed reduced nephrotoxicity in mice while retaining antibacterial activity.

Conclusions

Overall, this proof-of-concept study demonstrated that our SIR model of polymyxins with hPepT2 provides a viable approach for the discovery of novel, safer lipopeptide antibiotics.

Introduction

The World Health Organization has highlighted antimicrobial resistance as an urgent threat to global public health (1–3). In particular, the Gram-negative ‘superbugs’ Acinetobacter baumannii, Pseudomonas aeruginosa, and Klebsiella pneumoniae can develop resistance to nearly all available antibiotics, posing some of the most serious challenges in modern medicine (4). Due to the lack of effective alternative antibiotics, polymyxins are often used as a last-line therapy against these ‘superbugs’ (5). However, polymyxins can cause nephrotoxicity in up to 60% of patients after intravenous administration, which is the major dose-limiting factor limiting their clinical use (6, 7). Thus, there is an urgent need for safer, new-generation polymyxin lipopeptide antibiotics (8).

It has been demonstrated that polymyxin-induced nephrotoxicity is associated with extensive reabsorption into renal proximal tubular cells (6). The intracellular concentration of polymyxins in renal proximal tubular cells can be 5,000-fold higher than the extracellular concentration, which consequently leads to oxidative stress, autophagy, cell cycle arrest, and apoptosis (6). Therefore, the extensive reabsorption of polymyxins into renal proximal tubular cells is the very first step in nephrotoxicity.

Human Solute Carrier transporters (SLCs) are the primary influx transporters responsible for the cellular uptake of many drugs, including antibiotics (9, 10). There are over 300 SLC isoforms, among which the SLCO, SLC22A, and SLC15A subfamilies are key in mediating drug transport across the cell membrane (11, 12). Our previous study revealed that the cellular uptake of polymyxins is mediated through hPepT2 in the kidney (13), which is encoded by a key isoform of the SLC15A subfamily, human oligopeptide transporter 2 (hPepT2) (12). This transporter protein is highly expressed in renal tubular cells, mediating the uptake of polymyxins, with a comparable kinetic constant (Km, reflecting transporter-substrate binding affinity) to its prototype substrates (12–14). It is one of the proton-driven symporters, using the inwardly directed proton electrochemical gradient to drive the concentrative uptake of substrates across the cell membrane (15).

Substrates of hPepT2 include a range of di- or tri-peptides and peptide-like drugs, such as angiotensin-converting enzyme inhibitors (16). For instance, glycylsarcosine (Gly-Sar) is a synthetic dipeptide widely used as a model substrate for hPepT2, exhibiting a moderate binding affinity with a reported Km value of approximately 100 µM (17). Fluorescent peptide probes, such as 7-amino-4-methylcoumarin-3-acetic acid, have also been employed to evaluate hPepT2 function, with a reported Km value of approximately 780 µM (18). Several β-lactam antibiotics, including cefadroxil, have been identified as substrates of hPepT2, while antiviral agents such as valacyclovir can also be transported by this transporter (19, 20). Although many β-lactam antibiotics are substrates of hPepT2, some compounds, including cloxacillin and cefadroxil, have been reported to act as competitive inhibitors of hPepT2 (21). However, it remains unclear how hPepT2 recognises lipopeptide antibiotics, particularly polymyxins and this knowledge is critical for elucidating their interaction with polymyxins and its contribution to polymyxin-induced nephrotoxicity.

Here, we integrated computational and experimental approaches to elucidate the interaction between hPepT2 and polymyxins. Importantly, we conducted a proof-of-concept chemical biology study by designing novel lipopeptides and evaluating their cellular uptake, nephrotoxicity, and antibacterial activities. This study establishes the first structure-interaction relationship (SIR) model for polymyxins and hPepT2, laying the foundation for the rational drug design and discovery of novel lipopeptide antibiotics.

Results

MD prediction of the structural model of outward-facing hPepT2

We developed the three-dimensional structure of hPepT2 using AlphaFold2, as the outward structure of hPepT2 (Fig. 1a) has not been experimentally solved. The putative overall structure adopted an inward open conformation (Fig. 1b), which prevented the exploration of the initial capture of polymyxins from the extracellular side of the membrane. Parker et al. demonstrated the cryo-electron microscopy (cryo-EM) structure of rat PepT2 with an outward open conformation (22). As rat PepT2 shares 83% sequence similarity with hPepT2, it served as a suitable template for modelling the hPepT2 structure through homology modelling. Upon remodelling, the resultant structure of hPepT2 adopted an outward open conformation (Fig. 1c). Several transmembrane domains (TMs) are involved in the formation of the solvent-accessible cavity for polymyxin translocation, including TM1, TM2, TM5, TM7, TM8, and TM10. The gate residue pairs (L75-R329, P341-E79, D170-S635, and K642-A177) marking the conformations were clearly identified in the solvent-accessible regions on both the extracellular and intracellular surfaces (Fig. 1d). During the 500-ns all-atom MD simulations with the inward-facing conformation as an initial configuration, the distances between the extracellular gate residue pairs L75-R329 and P341-E79 increased from 20 Å to 31 Å, and from 15 Å to 21 Å, respectively. Meanwhile, the distances between the intracellular gate residue pairs D170-S635 and K642-A177 decreased from 15 Å to 13 Å, and from 13 Å to 10 Å, respectively (Fig. 1e). Notably, the transition from the inward conformation to the outward conformation uncovered by MD simulations is consistent with the homology model of rat PepT2 regarding the distance of the extracellular and intracellular gates (22). This observation indicates the accuracy of the proposed hPepT2 model.

Structural conformations of hPepT2.

(a) The structural topology of hPepT2. (b) The structural model of hPepT2 built with the AlphaFold approach. (c) The structural model of hPepT2 built with the homologous modeling method. The three-dimensional structure of rat PepT2 (PDB code: 7NQK) is used as the structural template. (d) The residues that are used to define the geometry of the channel of hPepT2. (e) The distances between the selected residues and representative structural snapshots are analyzed based on MD simulations.

Polymyxin B binds to the lateral opening site of hPepT2 to enter the transport pathway

To further explore how hPepT2 recognises polymyxin molecules, we examined their interaction through long-timescale coarse-grained MD simulations. Although ten polymyxin B molecules were introduced into the simulation system, only one polymyxin molecule was observed to bind to the entrance of the hPepT2 binding pocket, while the other polymyxin molecules attached to the membrane lipids (Fig. S1). Importantly, we discovered a distinct pathway by which polymyxins reach the extracellular lateral opening gate of hPepT2, as revealed by the simulation replicates (Fig. 2a and S2), with distinct equilibrated binding conformations (Fig. 2b). Our MD simulations predicted a possible transport pathway for polymyxins via hPepT2. Initially, polymyxin B attached to the plateau of the TM1-TM6 helix bundle, mainly driven by electrostatic interactions between the Dab5, Dab8 and Dab9 residues of polymyxin B and the E79, D208, E214 and D215 residues of hPepT2 (Fig. 3). Subsequently, polymyxin B translocated to the junction between the TM7-TM12 helix bundle and the extracellular domains (ECDs) of hPepT2, where the Dab1, Dab3 and Dab5 residues of polymyxin B electrostatically interacted with the D342 and E555 residues of hPepT2. Polymyxin B then stabilised at the lateral opening gate of hPepT2 (i.e., lateral entry site), where the Dab8 and Dab9 moieties of polymyxin B electrostatically interacted with the D342 and E555 residues of hPepT2. Moreover, the hydrophobic atoms in Dab8 and Dab9 interacted with the F200, I201 and M204 residues of hPepT2, while the fatty acyl group of polymyxin B bound to the hydrocarbon tails of the phospholipid bilayer.

Spontaneous interactions of polymyxin B with hPepT2.

(a) The interaction pathway of polymyxin B with hPepT2. The spatial locations of polymyxin B molecules are represented with colored spheres. The color spectrum from cyan to red indicates the simulation continuity from 0 to 3 μs. (b) The equilibrated conformation of the hPepT2-polymyxin complex according to the proposed pathway. The polymyxin B molecule is shown in brown and blue spheres with its D-Phe6, and N-terminal fatty acyl group labeled.

Key structural moieties and residues involved in the interaction of polymyxin B with hPepT2.

The surface potential of hPepT2 is shown with red color indicating negative potential and blue color indicating positive potential.

To examine the detailed interactions between hPepT2 and polymyxin B, we performed all-atom MD simulations of the hPepT2-polymyxin B complex. During system preparation, the polymyxin B molecule was placed in four different poses within the complex to enhance the sampling of binding dynamics. The results revealed that Dab1, Dab8, and Dab9 form strong electrostatic interactions with hPepT2, with interaction energies ranging from -180 to -260 kJ/mol (Fig. 4), while Dab3 and Dab5 may also contribute significantly. Together with observations from the coarse-grained simulations, these results indicate that the positively charged Dab residues are key determinants of their binding affinity to hPepT2.

Interaction energy between hPepT2 and each residue of polymyxin B.

The interaction energy was calculated using four MD simulation replicates. Note: Coul: electrostatic interaction; VDW: hydrophobic interaction.

Functional mutagenesis elaborates on the predicted transport pathway

At physiological pH, there are five positively charged residues (i.e. Dab residues) on the polymyxin molecule, and we assumed that the key negatively charged residues of hPepT2 identified by the MD simulations (i.e. E79, D208, E214, D215, D317, D342, E555, and E622) may play crucial roles in the recognition of and interactions with polymyxins. We employed functional mutagenesis to confirm their roles in the uptake of polymyxins, as predicted by our MD studies. Through the commonly used alanine-scanning mutagenesis approach, we constructed the corresponding mutants of hPepT2. Notably, alanine substitution not only impacted the side-chain length of the proposed residues but also altered their charges from negative to neutral. It is postulated that such a drastic change in the residues of hPepT2 may substantially affect their interactions with polymyxins.

The transport activity of these hPepT2 mutants was evaluated by assessing the uptake of H3-Gly-Sar (a typical substrate of hPepT2) (Fig. 5a) (13) as well as the fluorescent polymyxin probe MIPS-9541 (Fig. 5b). The fluorescence images were also captured as a qualitative measurement of the overexpressing cells exposed to our dansyl-fluorescent polymyxin probe MIPS-9541, which indicated consistent results (Fig. 6). Our data showed that E79A, D80A, and D208A mutants maintained transport activity in the uptake of both substrates (Fig. 5). In contrast, E214A, D215A, D342A, and E622A mutants had impaired function in the transport of either substrate. The uptake of Gly-Sar was reduced to 65.4 ± 15.8%, 42.7 ± 13.2%, 49.4 ± 14.2% and 6.4 ± 4.7% of the wild type, respectively (Fig. 5a). In the case of MIPS-9541, the cellular uptake was decreased to 67.6 ± 14.2%, 38.9 ± 12.2%, 53.7 ± 16.2%, and 14.3 ± 8.0% of the wild type, respectively. Interestingly, D317A and E555A showed differential activity in the uptake of these substrates (33.1 ± 12.7% of the control for D317A and 93.6 ± 14.4% of the control for E555A in the uptake of Gly-Sar; 94.9 ± 13.3% of the control for D317A and 47.7 ± 16.2% of the control for E555A in the uptake of MIPS-9541), suggesting the distinct roles of both residues in interactions with different substrates.

Uptake of 3H-Glycosarcosine and polymyxin B fluorescence probe MIPS-9541 by hPepT2 and its mutants.

HEK293 cells were transfected with hPepT2 and its mutant constructs. (a) Uptake of [3H]-Gly-Sar (5 µM) by hPepT2 or its mutants. (b) Uptake of MIPS-9541 (10 µM) by hPepT2 or its mutants. Data are presented as mean ± SD. All experiments were performed in triplicate independently. ***p < 0.001 vs. control by Welch’s t-test.

Uptake of the polymyxin B fluorescence probe MIPS-9541 in HEK293 cells overexpressing hPepT2 or its mutants assessed by fluorescence imaging.

Data are presented as mean ± SD. All experiments were performed in triplicate independently. Note: WT: wild type; NC: negative control.

Subsequently, kinetic analysis was conducted to characterise those hPepT2 mutants with impaired activity in the uptake of MIPS-9541 (i.e. E214A, D215A, D342A, and E555A). Notably, the E622A mutant was not included in this analysis due to its critically low transporter function. As shown in Table 1, the D215A mutant exhibited an approximately two-fold increase in both Km (p=0.033, 95% confidence interval of 7.5 to 91.5 µM) and Vmax (p=0.017, 95% confidence interval of 1510 to 7206 pmol/(µg/4 min)). It was also evident that the D342A mutant showed a mildly increased Km (66.9 ± 13.2 µM of D342A vs. 47.1 ± 11.9 µM of the wild type, p=0.13, 95% confidence interval of -9.1 to 48.7 µM) without altering its Vmax. The Vmax values of E214A and E555A were significantly reduced (2332 ± 24.2 pmol/(µg/4 min) of E214A, p=0.036, and 2376 ± 320.8 pmol/(µg/4 min) of E555A, p=0.022, vs. 4252 ± 652.9 pmol/(µg/4 min) of the wild type). However, the Km values of E214A and E555A mutants were moderately decreased compared to those of the wild type (24.4 ± 4.5 µM of E214A, p=0.07 and 26.7 ± 7.0 µM of E555A, p=0.08 vs. 47.1 ± 11.9 µM of the wild type).

Notably, a change in Vmax value may be due to altered transporter protein expression and/or transporter turnover rate. Thus, the total cell and plasma membrane expression of hPepT2 and its mutants were investigated to further examine the molecular mechanism underpinning the functional alterations of these mutants. As shown in Fig. 7, E214A, D215A, E555A, and E622A significantly reduced total cell expression and cell surface expression. In the case of D342A, it retained transporter protein expression in the whole cell but had a markedly impaired cell surface expression.

Total cell and cell surface expression of hPepT2-c-Flag and its mutants.

HEK293 cells were transfected with hPepT2-c-Flag and its mutant constructs. (a) Cells were lysed and subjected to SDS-electrophoresis. Immunoblots were probed with anti-Flag antibody followed by probing with an anti-actin antibody as a loading control. Representative images of each hPepT2 mutant are shown. (b) Cell surface proteins were labelled with NHS-ss-Biotin and then pulled down with Streptavidin-agarose beads. Biotin-labelled surface protein samples were subjected to SDS-electrophoresis and the immunoblots were probed with anti-Flag antibody. Equal portions of the supernatant samples were separated by SDS-PAGE, and the immunoblots were incubated with anti-actin antibody. Representative images of each hPeT2 mutant are shown. (c) Densitometry analysis of the relative total cell expression of hPepT2 mutants (ratios of Flag/actin). (d) Densitometry analysis of the relative cell surface expression of hPepT2 mutants (ratios of Flag/actin). Data are presented as % of the hPepT2 wildtype control (mean ± SD). Experiments were repeated on three occasions. *p < 0.05; **p < 0.01 vs. wild-type control by unpaired t-test. Note: WT: wild type; NC: vector-transfected negative control.

Chemical biology of polymyxin interaction with hPepT2

As described above, the Dab1, Dab3, Dab5, Dab8, and Dab9 residues of polymyxin B were predicted to be involved in the interaction with hPepT2. To test this hypothesis, we designed and synthesised five polymyxin B analogues with alanine substitutions on these Dab positions. We performed a transport uptake assay to evaluate their interactions with hPepT2. As shown in Fig. 8, the lipopeptide analogues FADDI-170, FADDI-175, FADDI-793, and FADDI-795 but not FADDI-167, showed reduced uptake in hPepT2-expressing cells, indicating that Dab1, Dab3, Dab5 and Dab9 but not Dab8, are likely involved in the interaction with hPepT2. Therefore, these four lipopeptide analogues may have less intracellular accumulation and reduced nephrotoxicity.

Uptake of the polymyxin B fluorescence probe MIPS-9541 and polymyxin analogues by hPepT2.

The data are expressed as the fold change in the uptake via hPepT2 vs. that of the vector-transfected control. Data are presented as mean ± SD. All experiments were performed in triplicate independently. ***p < 0.001 vs. MIPS-9541 control by Welch’s t-test.

Pharmacological evaluations of lead polymyxin analogues

We firstly assessed the minimum inhibitory concentrations (MICs, the lowest concentration that inhibits visible growth of a microorganism after overnight culture) of the four lipopeptide analogues predicted to have disrupted interactions with hPepT2. As shown in Table 2, FADDI-170 showed significantly reduced activity against FADDI-PA025, FADDI-EC006, FADDI-EC003, FADDI-EC001, FADDI-AB034, as well as A. baumannii ATCC 19606 and ATCC 17978 (MICs increased ≥16-fold, compared to polymyxin B). FADDI-175 displayed low activity against all the tested strains (MICs ≥32 µg/mL), which was not further considered for subsequent testing. FADDI-793 was less active than polymyxin B against A. baumannii ATCC 19606, ATCC 17978, FADDI-AB034, FADDI-EC006, and FADDI-EC003. FADDI-795 was the most active among them, showing comparable MICs against these bacterial strains except for FADDI-PA025, when compared to polymyxin B. Interestingly, FADDI-793 and FADDI-795 were more active than polymyxin B against K. pneumoniae ATCC 13883. We then employed the mouse model established in our laboratory (23) to examine the nephrotoxicity of the four antibacterial polymyxin analogues. Interestingly, FADDI-795 showed no observable nephrotoxicity, while the other three analogues possessed mild kidney toxicity, comparable to that of polymyxin B (Table 2).

Overall, the transporter uptake study and pharmacological assessments of these polymyxin analogues provide critical mechanistic data on the interaction and transport of polymyxins by hPepT2. Importantly, our integrated computational and cell biology approach highlighted the feasibility of reducing polymyxin nephrotoxicity by disrupting their interactions with hPepT2. Moreover, hPepT2 is a viable target for designing novel polymyxin-like lipopeptide antibiotics with reduced nephrotoxicity.

Discussion

Life-threatening MDR bacterial infections rank among the world’s most pressing health threats (24), yet the discovery of antibiotics with novel mechanisms of action has stalled for decades. Polymyxins remain a last-line therapy for MDR Gram-negative bacterial pathogens, particularly in low- to middle-income countries; however, their clinical use has been significantly limited due to nephrotoxicity (5, 6).

Our previous study demonstrated that polymyxins are substrates of hPepT2, which mediates the substantial reabsorption of polymyxins in renal proximal tubular cells, leading to nephrotoxicity (13). However, little is known about the interaction between polymyxin molecules and hPepT2. Notably, previous studies have reported that megalin mediates the cellular transport of polymyxins in the kidney via a receptor-mediated endocytosis mechanism (6, 25). Receptor-mediated endocytosis is closely linked to changes in signalling pathways; thus, disrupting the interaction between polymyxins and megalin may lead to potential signalling consequences and complicate treatment outcomes. In contrast, transporter-mediated uptake does not involve such signalling changes, making manipulation of polymyxin-hPepT2 interactions a more favourable strategy for mitigating polymyxin-associated nephrotoxicity. Furthermore, our previous study demonstrated a direct structural interaction of polymyxins with the membrane of human kidney proximal tubular cells (26), showing that polymyxin B inserts its fatty acyl tail into the hydrophobic region of the cell membrane. The stereochemistry at position 3 and the hydrophobicity at positions 6 and 7 of polymyxin B were found to be critical for this interaction. These findings provide important insights into how polymyxins alter membrane dynamics and lipid organization in kidney cells, contributing to a better understanding of polymyxin-induced nephrotoxicity.

The current study developed the first SIR model of polymyxins with hPepT2, expediting the discovery of novel safer lipopeptide antibiotics. We firstly employed MD simulations to predict the binding of polymyxins to hPepT2 (Fig. 1 and 2) and identified several critical amino acid residues of hPepT2 in the recognition of polymyxins. An outward-facing model of hPepT2 was first constructed, indicating the access of polymyxin molecules to the lateral opening gate of hPepT2 (Fig. 3). Limited by the sampling capability of MD simulations, the complete translocation pathway of polymyxin B through hPepT2 was not observed. The distinct conformation of polymyxin B indicates that it is situated in the central binding cavity of hPepT2. To further explore the conformational change of polymyxin molecules required for hPepT2 transport, all-atom MD simulations were conducted by varying the orientations of polymyxin B. Our results revealed that this conformational transition involved a notable flip of polymyxin B, which enhanced the interaction between its fatty acyl chain and the deeper regions of the predicted hPepT2 binding pocket. Notably, the subsequent transporter mutagenesis and molecular characterisations of hPepT2 explored our proposed SIR model. Our integrated computational and experimental findings provide a proof-of-concept that disrupting the interaction between polymyxin and hPepT2 is a novel strategy to reduce their nephrotoxicity.

As polymyxins are cationic at physiological pH, we strategically selected the negatively charged amino acids of hPepT2 (i.e. aspartic acid and glutamic acid) for further investigation. The significantly reduced uptake of polymyxin B by hPepT2 mutants E214A, D215A, D342A, E555A, and E622A demonstrated the critical role of these negatively charged residues in transporting polymyxins (Fig. 5 and 6). Moreover, the kinetic results of these transporter mutants elucidated the mechanism underpinning such functional impairment (Table 1). As both transporter cell surface expression and turnover rate account for the maximum velocity Vmax of a transporter protein (12), it was critical to evaluate the protein expression of these transporter mutants in total cell lysates and at the cell surface (Fig. 7). Both the kinetic analysis and protein expression profiles (Table 3) showed different mechanisms of impaired polymyxin uptake by hPepT2. For example, it is highly likely that E214A reduced Vmax due to reduced total cell and membrane protein expression. D215A reduced polymyxin uptake mainly due to impaired binding affinity to polymyxins (i.e. increased Km). The transporter turnover rate of D215A might increase to compensate for the reduced cell surface expression; thus, its Vmax was overall increased. Similarly, the decreased polymyxin uptake mediated via D342A may be primarily attributed to its reduced polymyxin binding affinity. Its decreased membrane protein expression and likely elevated transporter turnover rate may have resulted in an unchanged Vmax. In contrast, the dysfunction of E555A may be mainly due to its reduced Vmax, likely resulting from a reduced total cell and membrane protein expression, although its binding affinity to polymyxins is slightly increased (i.e. decreased Km). The kinetic parameters of E622A were unable to be determined due to its low polymyxin uptake. Our data also indicated that its low protein expression in total cell and at cell surface may contribute to its impaired transport activity in polymyxin uptake. Notably, reduced total cellular expression of membrane transporters may result from impaired protein stability (27). In addition, cell surface transporter proteins undergo constitutive subcellular trafficking processes, including internalisation, recycling, and membrane targeting (28). Therefore, decreased cell surface expression of transporter proteins may arise from reduced total cellular expression and/or disrupted subcellular trafficking (27). Our findings suggest that E214, D215, E555, and E622 residues may play important roles in maintaining hPepT2 protein stability, whereas substitution of these residues with alanine may influence the subcellular trafficking of hPepT2. Nevertheless, the detailed molecular mechanisms underlying the changes in hPepT2 protein expression associated with these residues warrant further investigation but are beyond the scope of the current study.

Notably, because some transporter mutants (e.g., D342A) exhibited both increased Kmand reduced membrane expression, it is not possible to conclusively determine whether the decreased transporter function results from reduced binding affinity, lower protein expression, or a combination of both. This is a limitation of the current study. Nevertheless, our molecular assay results support the applicability of the bioinformatic predictions, as they allow us to focus the analysis on a limited set of transporter residues most relevant to our SIR model.

To validate the MD predictions, we further investigated the transport function of these mutant transporters through their uptake of the classic hPepT2 substrate Gly-Sar (Fig. 5A). The uptake of Gly-Sar was dramatically decreased in the mutants E214A, D215A, D317A, D342A, and E622A, which is consistent with the reduced cell surface expression of E214A, D215A, D342A, and E622A (Fig. 7). D317A is exceptional as it showed decreased uptake of Gly-Sar but maintained transport function in the case of polymyxin B, which indicated that the residue D317 may be only important for the uptake of Gly-Sar but not polymyxins. E555A is another exception, which remained functional for the uptake of Gly-Sar but reduced uptake of polymyxins. As E555A had a reduced total cell and cell surface expression, its increased apparent affinity for Gly-Sar may partially compensate for the reduced Vmax. Thus, the binding determinants of polymyxins may not fully overlap with those of Gly-Sar. Therefore, the transport inhibition study was only employed as a positive control initially and not considered in the subsequent evaluation of the antibacterial activity of polymyxin analogues.

According to the predicted topology model of hPepT2 (12), E214A and D215A are situated at the ECD between TM5 and TM6. D342A resides at the ECD between TM7 and TM8. E555A is located at the ECD between TM9 and TM10. E622A is situated within TM10. It was reported that the large ECD between TM9 and TM10 interacts with TM1 to maintain transport function (29). In our study, we demonstrated that substitution of E555 mildly increased the apparent binding affinity for polymyxin B while reducing Vmax, likely resulting from decreased protein expression. These findings highlight the need for more in-depth mechanistic understanding of hPepT2-mediated peptide transport. Notably, E622A in TM10 has a very low transport function in the uptake of polymyxin B, which is largely due to its low protein expression in the total cell and on the cell surface (Fig. 5 and 7). Therefore, this residue appears to be essential for preserving hPepT2 stability. Similar findings have been reported in other SLCs (27, 30, 31). According to the protein sequence alignment between hPepT2 and another key isoform of the SLC15A subfamily, hPepT1, E595 of hPepT1 is the equivalent residue to E622 in hPepT2. It has been shown that the alteration of E595 residue to cysteine substantially impacted transport function but not protein expression (32). This finding reflects the differing substrate specificities and transport capacities of hPepT1 and hPepT2 (12), highlighting the essential role of hPepT2 in mediating polymyxin transport.

Upon validating the predictions of the SIR model of polymyxins and hPepT2, our study was advanced to employ a chemical biology approach to rationally design and in-house synthesize polymyxin analogues. With the hope of disrupting the interaction between hPepT2 and these analogues, we aimed to discover candidate lipopeptide antibiotics with reduced nephrotoxicity while maintaining similar or better antibacterial activity. Based on the SIR model, the Dab1, Dab3, Dab5, Dab8, and Dab9 residues of polymyxins are likely to be involved in their interaction with hPepT2 (Fig. 4). Thus, we synthesised FADDI-167, FADDI-170, FADDI-175, FADDI-793, and FADDI-795 with alanine substitutions at these Dab residues. We first examined the hPepT2-mediated uptake of these polymyxin analogues. Interestingly, we discovered that all analogues except FADDI-167 disrupted the uptake via hPepT2 (Fig. 8). This finding indicated that FADDI-170, FADDI-175, FADDI-793 and FADDI-795 possibly impaired their uptake via hPepT2, which may contribute to reduced renal accumulation and nephrotoxicity. Thus, these four analogues were subjected to further pharmacological evaluations. Interestingly, FADDI-795 and FADDI-175 showed the best and least antibacterial activity among them, respectively (Table 4). Finally, we employed our mouse model to evaluate the nephrotoxicity of these lipopeptides. Using histological evaluations, FADDI-795 showed no observable renal toxicity in this mouse model and emerged as the lead compound with comparable antibacterial activity. A limitation of this research is that the nephrotoxicity assessment of new lipopeptides was performed in only a small number of mice in our proof-of-concept study. Nevertheless, the findings indicate that FADDI-795 is a promising candidate that will undergo further pharmacological evaluations to explore its translational potential, such as comprehensive pharmacokinetic studies, renal accumulation assessment and efficacy tests in relevant infectious disease models.

Overall, our study developed the first SIR model of polymyxins with hPepT2 by integrating computational, chemical biology, and molecular approaches. Our molecular studies provide the first experimental validation of the predicted polymyxin-hPepT2 binding model. These mechanistic insights are essential for elucidating the mechanism of dose-limiting nephrotoxicity associated with polymyxin treatment. Moreover, our ongoing drug discovery efforts will further refine this SIR model and expedite the discovery and development of novel, safer lipopeptide antibiotics against Gram-negative bacterial infections.

Materials and Methods

Reagents and chemicals

[3H] Gly-Sar (2 Ci/µmol) was obtained from BioScientific Pty. Ltd. (Gymea, NSW, Australia). Dulbecco’s Modified Eagle Medium (DMEM), fetal bovine serum (FBS), and other culture supplements were purchased from Thermo Scientific (Lidcombe, NSW, Australia). Gly-Sar and polymyxin B1 for molecular studies were purchased from Sapphire Biosciences (Redfern, NSW, Australia). Polymyxin B sulfate for animal studies was purchased from BetaPharm (Shanghai, China). MIPS-9541 was synthesised as previously reported (13, 33, 34). Plasmids containing the full-length hPepT2 and hPepT2-c-Flag were obtained from Australian Biosearch (Balcatta, WA, Australia). All other reagents and chemicals were purchased from Sigma-Aldrich (Castle Hill, NSW, Australia).

Structural modelling of hPepT2 and polymyxins

The inward-open conformation of hPepT2 was predicted using AlphaFold2 via the ColabFold implementation with default parameters, employing the MMseqs2 multiple sequence alignment pipeline. To obtain the physiologically relevant outward-open conformation of hPepT2 for substrate binding, homology modeling was performed using MODELLER with the cryo-EM structure of rabbit PepT2 in the outward-open state (PDB: 7NQK) as the template. The sequence alignment between hPepT2 and rabbit PepT2 was conducted using Clustal Omega. One hundred models were generated, and the final model was selected based on the lowest Discrete Optimized Protein Energy score and verified by Ramachandran plot analysis (with >95% of residues in favored regions). The structure of polymyxin B1 was constructed and energy-minimized using the CHARMM36 force field. Computational alanine scanning of polymyxin B1 was performed by individually replacing each Dab (2,4-diaminobutyric acid) residue at positions 1, 3, 5, 8, and 9 with alanine using the mutagenesis wizard in PyMOL.

Coarse-grained molecular dynamics (MD) simulations

Polymyxin B1 was employed as the prototype (27) and the coarse-grained (CG) model system of hPepT2 was constructed using the Martini Maker module in CHARMM-GUI (28). The membrane lipid composition was 50% phosphatidylcholine, 25% phosphatidylethanolamine, and 25% phosphatidylserine (29). In each simulation system, 10 coarse-grained polymyxin molecules were randomly introduced to enhance the probability of binding to hPepT2. These were solvated using CG water particles and neutralised with 0.1 M sodium chloride in a 15 × 15 × 14 nm3 simulation box. All CG simulations were conducted under periodic boundary conditions using the GROMACS program (version 2021.2) and the Martini 2.2 force field (30, 31). To eliminate steric interference, independent steepest energy minimization was performed for every system to achieve the maximum force below 1,000 kJ mol-1 nm-2. Subsequently, a six-step equilibration cycle was carried out by gradually turning off the position restraints on the lipid and polymyxin molecules. Finally, 3 µs production simulations were performed for each of two independent replicates (including independent energy minimization) were conducted for each system in the NPT (referring to the number of particles (N), system pressure (P), and temperature (T)) ensemble. Two replicates per condition were performed at this exploratory stage, as the CG-MD simulations were designed to identify the binding pathway of polymyxin B to hPepT2; the final binding poses from both replicates were highly consistent. Quantitative interaction analysis was subsequently performed using all-atom MD simulations with four replicates. The temperature and pressure were coupled to 310 K using the velocity rescaling method (time constant of 1 ps) and 1 bar using the semi-isotropic barostat and Parrinello-Rahman algorithm (time constant of 5 ps), respectively (32). Molecular visualizations, structural figures, and trajectory animations were prepared using PyMOL (Schrodinger) and ChimeraX. An animation of the CG-MD binding trajectory was provided as Supplementary Movie S1.

All-atom MD simulations

The previous study of PepTXc, a homolog of hPepT2, revealed that the protonation of its His67 promotes the outward-opening conformation (33); therefore, the corresponding residue His87 of hPepT2 was set to be protonated during all-atom simulations to bias the system toward an outward opening state. In the two CG-MD replicates, polymyxin B eventually bound to the lateral opening gate of hPepT2. To examine the detailed interaction between polymyxin and hPepT2, a single polymyxin B1 molecule was docked into the predicted binding pocket of hPepT2 using AutoDock Vina (34). Furthermore, to enhance the sampling of the binding dynamics, four different poses of polymyxin B1 were employed to build the initial hPepT2-polymyxin complex. Force field parameters for polymyxin B1 were generated using CGenFF. The membrane protein system was then constructed using the membrane builder module in CHARMM-GUI (35). The membrane lipid composition was 50% phosphatidylcholine, 25% phosphatidylethanolamine, and 25% phosphatidylserine, consistent with the CG-MD simulations. Each system was solvated with TIP3P water molecules in a rectangular box of 11 × 11 × 14 nm3 and neutralized with 0.15 M NaCl, yielding a total of approximately 158,000 atoms.

GROMACS 2021.2 was used to perform all MD simulations with the CHARMM36 all-atom force field (36). A time step of 2 fs was employed to integrate the Newtonian equations of motion, and all bonds involving hydrogen atoms were constrained. Long-range electrostatic interactions were treated using the Particle Mesh Ewald (PME) method with a switching function applied between 10 and 12 Å (37). Energy minimization was performed using the steepest descent algorithm. Subsequently, the system was equilibrated through a series of MD simulation steps in the NVT (referring to the number of particles (N), system volume (V), and temperature (T)) and NPT ensembles (310 K, 1 bar), during which positional restraints on membrane and protein atoms were gradually reduced. The root mean square deviations of the hPepT2-polymyxin B1 complex were calculated to assess simulation equilibration (Fig. S3). No restraints were applied in the final stage of equilibration and throughout the production run. The PME method was employed to treat long-range electrostatic interactions with a short-range cut-off of 1.2 nm, while the shifted Lennard-Jones potential algorithm was used to calculate van der Waals interactions with a general cut-off of 1.2 nm and a shifting cut-off of 1.0 nm (38, 39). The trajectory during production simulations was recorded every 10 ps. Production simulations were conducted for 500 ns for outward-open conformation sampling and 100 ns for polymyxin-hPepT2 interaction analysis.

Site-directed mutagenesis and plasmid DNA sequencing

The hPepT2 mutants were generated using a QuickChange II site-directed mutagenesis kit (Agilent Technologies, Mulgrave, Victoria, Australia) with the primers listed in Table 4. All mutant construct sequences were confirmed by DNA sequencing (Ramaciotti Centre for Gene Function Analysis, Randwick, NSW, Australia) with the BigDyeTM Terminator v3.1 kit (Thermo Scientific, Lidcombe, NSW, Australia) (11, 27, 35–37).

Transfection of hPepT2 and its mutants into HEK293 cells

HEK293 cells were cultured in DMEM supplemented with 10% FBS at 37°C with 5% CO2. Cells were transfected with hPepT2 or its mutant plasmid DNA using Lipofectamine 2000 Reagent according to the manufacture’s instructions (Invitrogen, Mount Waverley, VIC, Australia) (9, 13, 28, 38, 39). At 24 h after transfection, the cellular uptake of [3H] Gly-Sar or MIPS-9541 was assessed.

Transport uptake assay

To evaluate the uptake of [3H] Gly-Sar via hPepT2 or its mutant transporters, we measured the accumulation of [3H] Gly-Sar in hPepT2-overexpressing HEK293 cells. The uptake was conducted at 37°C in phosphate-buffered saline (PBS, pH 5.0) containing 5 mM glucose. The specific substrate concentration was 2.5 μM [3H] Gly-Sar, and the duration was 8 min according to our previous studies (9, 13). The uptake assay was terminated by three rapid washes with ice-cold PBS. Samples were lysed in 0.2 M NaOH, neutralized in 0.2 M HCL and then processed for liquid scintillation counting. The uptake counts for hPepT2 or its mutant expressing cells were all subtracted by those of the control.

To evaluate the uptake of polymyxin B via hPepT2 or its mutant transporters, we measured the accumulation of a validated fluorescent polymyxin probe MIPS-9541 and polymyxin analogues (i.e. FADDI-167, -170, -175, -793, and -795, Fig. S4) in hPepT2-overexpressing HEK293 cells (13). The uptake was performed in PBS (pH 5.0) supplemented with 5 mM glucose for 10 min at 37°C. Intracellular fluorescence accumulation was quantified with excitation/emission wavelengths of 350 nm/518 nm using a Tecan Safire II microplate reader (Thermo Scientific, Lidcombe, NSW, Australia). Background counts of the vector-transfected cells were subtracted from all uptake measurements. Kinetic analysis was conducted by measuring the uptake of transporter-expressing cells with a range of concentrations of MIPS-9541 (0 - 50 µM) over 10 min. Apparent Km and Vmax values were determined using GraphPad Prism 10.0 (13, 40).

Fluorescence imaging of MIPS-9541

HEK293 cells were seeded on chamber slides and transfected with hPepT2 or its mutant plasmids. After 24 h, the culture medium was removed and cells were incubated with freshly prepared MIPS-9541 solution in PBS (10 µM, pH 5.0) for 10 min at 37°C. Cells were then washed with ice-cold PBS three times and mounted in SlowFade Gold Antifade Mountant reagent supplemented with 4’,6-diamidino-2-phenylindole (DAPI, Thermo Scientific, Lidcombe, NSW, Australia). Samples were imaged with a Leica Thunder 3D Imager (Leica Microsystems, North Ryde, NSW, Australia).

Cell surface biotinylation

We employed sulfo-NHS-SS-Biotin to label the plasma membrane proteins in HEK293 cells overexpressing hPepT2 or its mutants (41). Cell culture plates were pre-chilled on ice. Culture medium was aspirated, and cells were washed twice with cold PBS (pH 8.0). Freshly prepared 1 mg/mL sulfo-NHS-SS-Biotin in PBS (pH 8.0) was then added to each well and incubated for 30 min. Cells were washed twice with 100 mM glycine in PBS (pH 8.0) and then three times with PBS (pH 7.4). Cells were lysed in the lysis buffer (Tris 10 mM, NaCl 150 mM, EDTA 1 mM, SDS 0.1%, and Triton X-100 1% with 1:1,000 dilution of a protease inhibitor cocktail) (27). Cell lysate was subjected to centrifugation at 14,000 g and 4°C for 10 min, and the supernatant was collected. Pre-washed Streptavidin-Agarose beads were supplied to the samples and incubated with rotation at 4°C for 1 h. After three rounds of centrifugation and washing with PBS (pH 7.4), beads were resuspended in PBS with 1x Laemmli buffer containing 2.5% β-mercaptoethanol and incubated at 55°C for 30 min. The remaining supernatant was collected, denatured, and subjected to electrophoresis.

Total cell lysis, electrophoresis and immunoblotting

Upon washing, cells were collected and lysed in the lysis buffer. After incubating on ice for 10 min, cell lysate was centrifuged at 14,000 g and 4°C for 10 min. The supernatant was collected as total cell lysate and denatured at 55°C in 1x Laemmli buffer containing 2.5% β-mercaptoethanol for 30 min.

Protein lysate samples were loaded onto 10% SDS-PAGE gels followed by electrophoresis as described before (42), and the protein samples on the gels were transferred to polyvinylidene difluoride (PVDF) membranes. Following blocking with 5% non-fat milk in PBS-Tween (Na2HPO4 80 mM, KH2PO4 20 mM, NaCl 100 mM, and 0.05% Tween 20, pH 7.5) and multiple washings with PBS-Tween, the immunoblots were incubated at 4°C overnight with anti-flag (DYKDDDDK) antibody (1:1,000; Catalogue Number 2368, Genesearch, Arundel, QLD, Australia) or anti-β-actin antibody (1:2,000, Catalogue Number 4967, Genesearch). After washing with PBS-Tween three times, the immunoblots were incubated at room temperature for 1 h with goat anti-rabbit IgG conjugated with HRP (1:5,000; Catalogue Number A0545, Sigma-Aldrich) or goat anti-mouse IgG conjugated with HRP (1:10,000; Catalogue Number A2554, Sigma-Aldrich). The blots were washed, incubated with Immobilon Western Chemiluminescent HRP substrate (Merck Millipore, Kilsyth, VIC, Australia), and imaged with an ImageQuant LAS 500 (GE Healthcare, Avantor, PA, USA). The densitometry analysis of immunoblots was processed with ImageJ (NIH, Bethesda, MD, USA).

Chemical synthesis of polymyxin analogues

Five lipopeptides (FADDI-167, FADDI-170, FADDI-175, FADDI-793, and FADDI-795) were synthesised and purified in house as described before (23, 43). Briefly, the lipopeptides were prepared using a Protein Technologies Prelude automated peptide synthesizer according to the standard protocol for Fmoc solid-phase peptide chemistry.

Measurements of minimum inhibitory concentrations

MICs were measured using the broth micro-dilution method with clinical isolates and American Type Culture Collection (ATCC) strains (44). Each bacterial strain (∼106 colony-forming units [CFU] per mL) was suspended in 100CμL of freshly prepared cation-adjusted Mueller-Hinton broth (CaMHB) in 96-well plates. Lipopeptides were diluted into different concentrations in CaMHB and added to the plates for subsequent incubation at 37°C for 18-24 h. MICs were determined as the lowest concentration at which the visible growth of bacteria was inhibited.

Nephrotoxicity in mice

The mouse nephrotoxicity study was approved by the Monash Animal Ethics Committee. Mice were maintained in micro-isolators in a temperature-controlled PC2 animal laboratory with an ambient humidity of 50-70% and a 12Ch/12Ch dark/light cycle. Lipopeptide solutions were subcutaneously administered to mice with a dose of 8 mg/kg every 2 h for 6 doses in one day (a total dose of 48 mg/kg/day, n=3 each group). At 24 h after the last injection, mice were sacrificed, the kidneys were collected from mice immediately, and fixed in 10% formalin buffer (pH 7.4, Sigma-Aldrich, Australia). Histological examination of the kidney samples was conducted at the Australian Phenomics Network-Histopathology and Organ Pathology (University of Melbourne, Parkville, VIC, Australia). An experienced pathologist blinded to treatment groups assessed renal histology and graded the severity of nephrotoxicity using a semi-quantitative scoring system (scores 1-10) as previously described (23). The percentage of kidney affected was also scaled in the range of 0-6 (0: <1%; 1: 1 to <5%; 2: 5 to <10%; 3: 10 to <20%; 4: 20 to <30%; 5: 30 to ≤40% and 6: >40%). A semi-quantitative score (SQS) was calculated using the grade score and kidney damage percentage to quantify the overall kidney damage (23).

Statistical analyses

Unpaired t-test was employed to detect the variance between the protein expression of hPepT2 and its mutants. Welch’s t-test was used to assess the differences between the transport function or kinetic parameters of hPepT2 and its mutants. All data are presented as mean ± standard deviation (SD), and a p value <0.05 was considered statistically significant. All experiments were repeated three times with triplicates in each experiment.